\chapter{Some Basic Notions of Set Theory}

\begin{problembox}[2.1: Equality of Ordered Pairs]
Prove Theorem 2.2: $(a, b) = (c, d)$ if and only if $a=c$ and $b=d$.

Hint: $(a, b) = (c, d)$ means $\{\{a\}, \{a, b\}\} = \{\{c\}, \{c, d\}\}$. Now appeal to the definition of set equality.
\end{problembox}

\textbf{Solution:}  
We must prove that $(a, b) = (c, d)$ if and only if $a=c$ and $b=d$.
The Kuratowski definition of an ordered pair is $(x, y) = \{\{x\}, \{x, y\}\}$.
If $a=c$ and $b=d$, then $(a,b) = \{\{a\}, \{a,b\}\} = \{\{c\}, \{c,d\}\} = (c,d)$. This direction is straightforward.

For the other direction, assume $(a, b) = (c, d)$. This means the sets are equal:
\[ \{\{a\}, \{a, b\}\} = \{\{c\}, \{c, d\}\} \]
By the definition of set equality, each element of the first set must be an element of the second, and vice versa. We consider two cases.

\textbf{Case 1: $a=b$.}
In this case, $(a, a) = \{\{a\}, \{a, a\}\} = \{\{a\}\}$.
For the sets to be equal, we must have $\{\{c\}, \{c, d\}\} = \{\{a\}\}$. This implies that the set $\{\{c\}, \{c, d\}\}$ has only one element, which means $\{c\} = \{c, d\}$. This equality holds if and only if $c=d$.
So we have $\{\{c\}\} = \{\{a\}\}$, which implies $\{c\} = \{a\}$, and thus $c=a$.
Since $c=d$ and $c=a$ and we started with $a=b$, we conclude that $a = b = c = d$. In particular, $a=c$ and $b=d$.

\textbf{Case 2: $a \neq b$.}
In this case, the set $\{\{a\}, \{a, b\}\}$ contains two distinct elements: the set $\{a\}$ with one member, and the set $\{a, b\}$ with two members. Therefore, the set $\{\{c\}, \{c, d\}\}$ must also contain two distinct elements, which implies $c \neq d$.
Since the sets are equal, their elements must match. We have two possibilities:
\begin{enumerate}
\item $\{a\} = \{c\}$ and $\{a, b\} = \{c, d\}$.
From $\{a\} = \{c\}$, we get $a=c$. Substituting this into the second equality gives $\{a, b\} = \{a, d\}$. Since $a \neq b$, the set on the left has two distinct elements. For the sets to be equal, we must have $b=d$. Thus, $a=c$ and $b=d$.
\item $\{a\} = \{c, d\}$ and $\{a, b\} = \{c\}$.
The first equality, $\{a\} = \{c, d\}$, would mean that the set $\{a\}$, which has one element, is equal to the set $\{c, d\}$, which has two elements (since $c \neq d$). This is impossible.
\end{enumerate}
The only possibility is that $a=c$ and $b=d$. In both cases, the equality of ordered pairs implies the equality of their corresponding components

\begin{problembox}[2.2: Properties of Relations]
Determine which of the following relations $S$ on $\mathbb{R}^2$ are reflexive, symmetric, and transitive:
\begin{enumerate}[label=(\alph*)]
\item $S = \{(x,y) \in \mathbb{R}^2 : x \leq y\}$
\item $S = \{(x,y) \in \mathbb{R}^2 : x < y\}$
\item $S = \{(x,y) \in \mathbb{R}^2 : x > y\}$
\item $S = \{(x,y) \in \mathbb{R}^2 : x^2 + y^2 \geq 1\}$
\item $S = \{(x,y) \in \mathbb{R}^2 : x^2 + y^2 < 0\}$
\item $S = \{(x,y) \in \mathbb{R}^2 : x^2 + x \leq y^2 + y\}$
\end{enumerate}
\end{problembox}

\textbf{Solution:}  
\begin{enumerate}[label=(\alph*)]
\item \textbf{Reflexive:} Yes - for all $x \in \mathbb{R}$, we have $x \leq x$ \\
\textbf{Symmetric:} No - if $x \leq y$ and $x \neq y$, then $y \not\leq x$ \\
\textbf{Transitive:} Yes - if $x \leq y$ and $y \leq z$, then $x \leq z$

\item \textbf{Reflexive:} No - $x < x$ is never true \\
\textbf{Symmetric:} No - if $x < y$, then $y \not< x$ \\
\textbf{Transitive:} Yes - if $x < y$ and $y < z$, then $x < z$

\item \textbf{Reflexive:} No - $x > x$ is never true \\
\textbf{Symmetric:} No - if $x > y$, then $y \not> x$ \\
\textbf{Transitive:} Yes - if $x > y$ and $y > z$, then $x > z$

\item  \textbf{Reflexive:}] No - the condition $(x,x) \in S$ requires $2x^2 \geq 1$, which fails for any $x$ in the interval $(-\frac{1}{\sqrt{2}}, \frac{1}{\sqrt{2}})$. \\
\textbf{Symmetric:} Yes - if $x^2 + y^2 \geq 1$, then $y^2 + x^2 \geq 1$ due to the commutative property of addition. 
\textbf{Transitive:} No - a counterexample is needed. Let $x = 0.6$, $y = 0.9$, and $z = 0.6$.
\begin{itemize}
\item $(x,y) \in S$ because $(0.6)^2 + (0.9)^2 = 0.36 + 0.81 = 1.17 \geq 1$.
\item $(y,z) \in S$ because $(0.9)^2 + (0.6)^2 = 0.81 + 0.36 = 1.17 \geq 1$.
\end{itemize}
However, $(x,z) \notin S$ because $(0.6)^2 + (0.6)^2 = 0.36 + 0.36 = 0.72 < 1$.

\item \textbf{Reflexive:} No - $x^2 + x^2 = 2x^2 \geq 0$ for all $x \in \mathbb{R}$ \\
\textbf{Symmetric:} Yes - if $x^2 + y^2 < 0$, then $y^2 + x^2 < 0$ \\
\textbf{Transitive:} Vacuously true - the relation is empty

\item \textbf{Reflexive:} Yes - for all $x \in \mathbb{R}$, $x^2 + x \leq x^2 + x$ \\
\textbf{Symmetric:} No - if $x^2 + x \leq y^2 + y$ and $x \neq y$, then $y^2 + y \not\leq x^2 + x$ \\
\textbf{Transitive:} Yes - if $x^2 + x \leq y^2 + y$ and $y^2 + y \leq z^2 + z$, then $x^2 + x \leq z^2 + z$
\end{enumerate}

\begin{problembox}[2.3: Composition and Inversion of Functions]
The following functions \( F \) and \( G \) are defined for all real \( x \) by the equations given below. 

\textbf{Part 1.} In each case where the composite function \( G \circ F \) can be formed, give the domain of \( G \circ F \) and a formula (or formulas) for \( (G \circ F)(x) \):
\begin{itemize}
\item[(a)] \( F(x) = 1 - x \), \quad \( G(x) = x^2 + 2x \)
\item[(b)] \( F(x) = x + 5 \), \quad \( G(x) = \frac{|x|}{x} \), \( G(0) = 1 \)
\item[(c)] \( F(x) = \begin{cases} 2x, & 0 \le x \le 1 \\ 1, & \text{otherwise} \end{cases} \),  
\( G(x) = \begin{cases} 3x^2, & 0 \le x \le 1 \\ 5, & \text{otherwise} \end{cases} \)
\end{itemize}
\textbf{Part 2.} In the following, find \( F(x) \) if \( G(x) \) and \( G[F(x)] \) are given:
\begin{itemize}
\item[(d)] \( G(x) = x^3 \), \quad \( G[F(x)] = x^3 - 3x^2 + 3x - 1 \)
\item[(e)] \( G(x) = 3 + x + x^2 \), \quad \( G[F(x)] = x^2 - 3x + 5 \)
\end{itemize}
\end{problembox}

\textbf{Solution:}

\textbf{(a)}
The domain of both \(F\) and \(G\) is \(\mathbb{R}\), so the domain of \(G \circ F\) is \(\mathbb{R}\).
\begin{align*}
(G \circ F)(x) &= G(F(x)) \\
&= G(1-x) \\
&= (1-x)^2 + 2(1-x) \\
&= (1 - 2x + x^2) + (2 - 2x) \\
&= x^2 - 4x + 3
\end{align*}

\textbf{(b)}
The domain of \(F\) is \(\mathbb{R}\) and the domain of \(G\) is \(\mathbb{R}\), so the domain of \(G \circ F\) is \(\mathbb{R}\).
\begin{align*}
(G \circ F)(x) = G(F(x)) = G(x+5)
\end{align*}
We evaluate this based on the value of the input to \(G\), which is \(x+5\):
\[ (G \circ F)(x) = 
\begin{cases} 
\frac{|x+5|}{x+5} = 1, & \text{if } x+5 > 0 \implies x > -5 \\
1, & \text{if } x+5 = 0 \implies x = -5 \\
\frac{|x+5|}{x+5} = -1, & \text{if } x+5 < 0 \implies x < -5 
\end{cases}
\]
This simplifies to:
\[ (G \circ F)(x) = \begin{cases} -1, & x < -5 \\ 1, & x \ge -5 \end{cases} \]

\textbf{(c)}
The domain of \(G \circ F\) is \(\mathbb{R}\). We analyze the composition in pieces based on the definition of \(F(x)\).
\begin{itemize}
\item If $0 \le x \le 1$, then $F(x) = 2x$. The value of $F(x)$ is in the interval $[0, 2]$. We must check where $F(x)$ falls in the domain of $G$.
\begin{itemize}
\item If $0 \le F(x) \le 1$, which means $0 \le 2x \le 1$, or $0 \le x \le 0.5$. In this case, $G(F(x)) = 3(F(x))^2 = 3(2x)^2 = 12x^2$.
\item If $F(x) > 1$, which means $2x > 1$, or $0.5 < x \le 1$. In this case, $G(F(x)) = 5$.
\end{itemize}
\item If $x < 0$ or $x > 1$, then $F(x)=1$. Since this value is in the interval $[0,1]$, we use the first rule for $G$: $G(F(x)) = G(1) = 3(1)^2 = 3$.
\end{itemize}
Combining these results, we get the piecewise formula:
\[ (G \circ F)(x) = \begin{cases} 
3, & x < 0 \\
12x^2, & 0 \le x \le 0.5 \\
5, & 0.5 < x \le 1 \\
3, & x > 1
\end{cases}
\]

\textbf{(d)}
We are given $G(x) = x^3$ and $G[F(x)] = x^3 - 3x^2 + 3x - 1$.
The composition is $(F(x))^3$. We can recognize the expression for $G[F(x)]$ as the expansion of a cube:
\[ x^3 - 3x^2 + 3x - 1 = (x-1)^3 \]
Therefore, we have $(F(x))^3 = (x-1)^3$, which implies $F(x) = x-1$.

\textbf{(e)}
We are given $G(x) = 3 + x + x^2$ and $G[F(x)] = x^2 - 3x + 5$.
We set up the equation for the composition:
\begin{align*}
G(F(x)) &= 3 + F(x) + (F(x))^2 \\
x^2 - 3x + 5 &= 3 + F(x) + (F(x))^2
\end{align*}
Rearranging gives a quadratic equation in terms of $F(x)$:
\[ (F(x))^2 + F(x) + (3 - (x^2 - 3x + 5)) = 0 \]
\[ (F(x))^2 + F(x) + (-x^2 + 3x - 2) = 0 \]
We use the quadratic formula to solve for $F(x)$:
\begin{align*}
F(x) &= \frac{-1 \pm \sqrt{1^2 - 4(1)(-x^2+3x-2)}}{2(1)} \\
&= \frac{-1 \pm \sqrt{1 + 4x^2 - 12x + 8}}{2} \\
&= \frac{-1 \pm \sqrt{4x^2 - 12x + 9}}{2}
\end{align*}
The term under the square root is a perfect square: $4x^2 - 12x + 9 = (2x-3)^2$.
\[ F(x) = \frac{-1 \pm \sqrt{(2x-3)^2}}{2} = \frac{-1 \pm (2x-3)}{2} \]
This yields two possible functions for $F(x)$:
\begin{enumerate}
\item $F_1(x) = \frac{-1 + (2x-3)}{2} = \frac{2x-4}{2} = x-2$
\item $F_2(x) = \frac{-1 - (2x-3)}{2} = \frac{-2x+2}{2} = 1-x$
\end{enumerate}


\begin{problembox}[2.4: Associativity of Function Composition]
Given three functions \( F, G, H \), what restrictions must be placed on their domains so that the following four composite functions can be defined?
\[
G \circ F, \quad H \circ G, \quad H \circ (G \circ F), \quad (H \circ G) \circ F
\]
Assuming that \( H \circ (G \circ F) \) and \( (H \circ G) \circ F \) can be defined, prove the associative law:
\[
H \circ (G \circ F) = (H \circ G) \circ F
\]
\end{problembox}

\textbf{Solution:}  
To define \( G \circ F \), the range of \( F \) must be contained in the domain of \( G \).  
To define \( H \circ G \), the range of \( G \) must be contained in the domain of \( H \).  
Under these conditions,  
\[
(H \circ (G \circ F))(x) = H(G(F(x))) = ((H \circ G) \circ F)(x)
\]  
So function composition is associative wherever defined.

\begin{problembox}[2.5: Set-Theoretic Identities]
Prove the following set-theoretic identities:
\begin{itemize}
\item[(a)] \( A \cup (B \cup C) = (A \cup B) \cup C \), \quad \( A \cap (B \cap C) = (A \cap B) \cap C \)
\item[(b)] \( A \cap (B \cup C) = (A \cap B) \cup (A \cap C) \)
\item[(c)] \( (A \cap B) \cup (A \cap C) = A \cap (B \cup C) \)
\item[(d)] \( (A \cup B)(B \cup C)(C \cup A) = (A \cap B) \cup (A \cap C) \cup (B \cap C) \)
\item[(e)] \( A \cap (B - C) = (A \cap B) - (A \cap C) \)
\item[(f)] \( (A - C) \cap (B - C) = (A \cap B) - C \)
\item[(g)] \( (A - B) \cup B = A \) if and only if \( B \subseteq A \)
\end{itemize}
\end{problembox}

\textbf{Solution:}  
Each identity can be proven by element-chasing: assuming \( x \in \) one side and showing \( x \in \) the other side.  
For example, for (b), if \( x \in A \cap (B \cup C) \), then \( x \in A \) and \( x \in B \cup C \Rightarrow x \in (A \cap B) \cup (A \cap C) \). Similar for the reverse.

\begin{problembox}[2.6: Image of Unions and Intersections]
Let \( f: S \to T \) be a function. If \( A \) and \( B \subseteq S \), prove:
\[
f(A \cup B) = f(A) \cup f(B), \quad f(A \cap B) \subseteq f(A) \cap f(B)
\]
Generalize to arbitrary unions and intersections.
\end{problembox}

\textbf{Solution:}  
For any \( x \in A \cup B \), \( f(x) \in f(A) \cup f(B) \).  
For intersections, \( x \in A \cap B \Rightarrow f(x) \in f(A) \cap f(B) \), but the converse need not hold.  
Generalization:  
\[
f\left( \bigcup_i A_i \right) = \bigcup_i f(A_i), \quad f\left( \bigcap_i A_i \right) \subseteq \bigcap_i f(A_i)
\]

\begin{problembox}[2.7: Inverse Image Laws]
Let \( f: S \to T \), and for any \( Y \subseteq T \), define the inverse image:
\[
f^{-1}(Y) = \{x \in S \mid f(x) \in Y \}
\]
Prove:
\begin{itemize}
\item[(a)] \( f^{-1}(Y_1 \cup Y_2) = f^{-1}(Y_1) \cup f^{-1}(Y_2) \)
\item[(b)] \( f^{-1}(T - Y) = S - f^{-1}(Y) \)
\end{itemize}
Generalize to arbitrary unions and intersections.
\end{problembox}

\textbf{Solution:}  
(a) If \( x \in f^{-1}(Y_1 \cup Y_2) \), then \( f(x) \in Y_1 \cup Y_2 \Rightarrow x \in f^{-1}(Y_1) \cup f^{-1}(Y_2) \), and vice versa.  
(b) \( f(x) \notin Y \iff x \notin f^{-1}(Y) \Rightarrow x \in S - f^{-1}(Y) \)

\begin{problembox}[2.8: Image of Preimage and Surjectivity]
Prove that \( f[f^{-1}(Y)] = Y \) for every \( Y \subseteq T \) if and only if \( f \) is surjective.
\end{problembox}

\textbf{Solution:}  
If \( f \) is surjective, every \( y \in Y \) has a preimage in \( S \), so is included in \( f[f^{-1}(Y)] \).  
If \( f \) is not surjective, then some \( y \notin f(S) \), and so not in the image of any preimage — thus excluded from \( f[f^{-1}(Y)] \).

\begin{problembox}[2.9: Equivalent Conditions for Injectivity]
Let \( f: S \to T \) be a function. Show the following are equivalent:
\begin{itemize}
\item[(a)] \( f \) is injective
\item[(b)] \( f(A \cap B) = f(A) \cap f(B) \) for all \( A, B \subseteq S \)
\item[(c)] \( f^{-1}[f(A)] = A \) for all \( A \subseteq S \)
\item[(d)] For disjoint sets \( A, B \subseteq S \), \( f(A) \cap f(B) = \emptyset \)
\item[(e)] If \( B \subseteq A \), then \( f(A - B) = f(A) - f(B) \)
\end{itemize}
\end{problembox}

\textbf{Solution:}  
Each condition implies the others under the assumption that \( f(x_1) = f(x_2) \Rightarrow x_1 = x_2 \).  
E.g., (c) implies \( f^{-1}[f(\{x\})] = \{x\} \Rightarrow \) only one \( x \) maps to any \( f(x) \).

\begin{problembox}[2.10: Subset Transitivity]
Prove that if \( A \subseteq B \) and \( B \subseteq C \), then \( A \subseteq C \).
\end{problembox}

\textbf{Solution:}  
If \( x \in A \), then since \( A \subseteq B \), we have \( x \in B \), and since \( B \subseteq C \), we get \( x \in C \).  
Thus, every element of \( A \) is in \( C \), so \( A \subseteq C \).

\begin{problembox}[2.11: Finite Set Bijection Implies Equal Size]
If \( \{1, 2, \ldots, n\} \sim \{1, 2, \ldots, m\} \), prove that \( m = n \).
\end{problembox}

\textbf{Solution:}  
A bijection between two finite sets implies they have the same number of elements.  
So if such a bijection exists, then \( \#\{1, \ldots, n\} = n = m = \#\{1, \ldots, m\} \), hence \( n = m \).

\begin{problembox}[2.12: Infinite Sets Contain Countable Subsets]
If \( S \) is an infinite set, prove that \( S \) contains a countably infinite subset.
\end{problembox}

\textbf{Solution:}  
We can construct an injection from \( \mathbb{N} \) into \( S \):  
Select \( a_1 \in S \), then pick \( a_2 \in S \setminus \{a_1\} \), then \( a_3 \in S \setminus \{a_1, a_2\} \), and so on.  
Since \( S \) is infinite, this process never terminates. Thus, \( \{a_1, a_2, \ldots\} \subseteq S \) is countably infinite.

\begin{problembox}[2.13: Infinite Set Similar to a Proper Subset]
Prove that every infinite set \( S \) contains a proper subset similar (i.e., bijective) to \( S \) itself.
\end{problembox}

\textbf{Solution:}  
Let $S$ be an infinite set. By the result of Exercise 2.12, $S$ contains a countably infinite subset. Let this subset be $A = \{a_1, a_2, a_3, \dots \}$.
Let $S' = S \setminus A$ be the set of elements in $S$ but not in $A$. Then $S = A \cup S'$, and this union is disjoint.

We want to find a proper subset $T \subset S$ and a bijection $f: S \to T$.
Let's define the proper subset as $T = S \setminus \{a_1\}$. Clearly, $T$ is a proper subset of $S$ because it's missing the element $a_1$.

Now, we define a function $f: S \to T$ as follows:
\begin{itemize}
\item For any element $x \in S'$ (i.e., any element not in our countable subset $A$), we define $f(x) = x$.
\item For any element $a_n \in A$ (where $n$ is a positive integer), we define $f(a_n) = a_{n+1}$.
\end{itemize}
The domain of $f$ is $S' \cup A = S$. The range of $f$ is $S' \cup \{a_2, a_3, a_4, \dots\}$, which is exactly the set $S \setminus \{a_1\} = T$.

To prove that $f$ is a bijection, we must show it is both injective and surjective.
\begin{itemize}
\item \textbf{Injectivity:} Suppose $f(x_1) = f(x_2)$.
\begin{itemize}
\item If $f(x_1)$ is in $S'$, then $f(x_1)=x_1$ and $f(x_2)=x_2$, so $x_1=x_2$.
\item If $f(x_1)$ is in $\{a_2, a_3, \dots\}$, say $f(x_1) = a_{k+1}$, then both $x_1$ and $x_2$ must be elements from $A$. Specifically, $x_1 = a_k$ and $x_2 = a_k$. Thus $x_1=x_2$.
\end{itemize}
In all cases, $f(x_1)=f(x_2)$ implies $x_1=x_2$.
\item \textbf{Surjectivity:} Let $y$ be any element in the codomain $T = S \setminus \{a_1\}$.
\begin{itemize}
\item If $y \in S'$, then $f(y) = y$.
\item If $y \in \{a_2, a_3, \dots\}$, then $y=a_k$ for some $k \geq 2$. The element $x=a_{k-1}$ is in $S$ and $f(x) = f(a_{k-1}) = a_k = y$.
\end{itemize}
Every element in $T$ has a preimage in $S$.
\end{itemize}
Since $f$ is a bijection from $S$ to its proper subset $T$, the set $S$ is similar to a proper subset of itself.


\begin{problembox}[2.14: Removing Countable from Uncountable]
If \( A \) is a countable set and \( B \) an uncountable set, prove that \( B - A \sim B \).
\end{problembox}

\textbf{Solution:}  
Since \( A \) is countable and \( B \) is uncountable, \( B - A \) is uncountable.  
Also, \( A \cup (B - A) = B \). Define a bijection \( f \) from \( B \) to \( B - A \cup \{a_0\} \subset B \) by remapping countably many points.  
Thus, \( B \sim B - A \).

\begin{problembox}[2.15: Algebraic Numbers are Countable]
A real number is called \emph{algebraic} if it is a root of a polynomial with integer coefficients.  
Prove that the set of all polynomials with integer coefficients is countable, and deduce that the set of algebraic numbers is also countable.
\end{problembox}

\textbf{Solution:}  
Each polynomial can be represented by a finite tuple of integers (its coefficients). The set of finite sequences of integers is countable (a countable union of countable sets).  
Each polynomial has finitely many roots, so the set of all algebraic numbers is a countable union of finite sets → countable.


\begin{problembox}[2.16: Power Set of Finite Set]
Let \( S \) be a finite set with \( n \) elements, and let \( T \) be the collection of all subsets of \( S \).  
Show that \( T \) is finite, and determine how many elements it contains.
\end{problembox}

\textbf{Solution:}  
Each element of \( S \) may either be in or not in a subset.  
So the number of subsets is \( 2^n \). Hence \( \#T = 2^n \), and \( T \) is finite.

\begin{problembox}[2.17: Real Functions vs Real Numbers]
Let \( R \) be the set of real numbers and \( S \) the set of all real-valued functions with domain \( R \).  
Show that \( S \) and \( R \) are not equinumerous.
\end{problembox}

\textbf{Solution:}  
Assume toward contradiction that there is a bijection \( f: R \to S \).  
Define a function \( h(x) = f(x)(x) + 1 \). Then \( h \in S \), but there is no \( x \in R \) such that \( f(x) = h \), since \( f(x)(x) \ne h(x) \).  
Contradiction → no such bijection. Thus, \( S \) has strictly greater cardinality than \( R \).

\begin{problembox}[2.18: Binary Sequences are Uncountable]
Let \( S \) be the set of all infinite sequences of 0s and 1s. Show that \( S \) is uncountable.
\end{problembox}

\textbf{Solution:}  
Use Cantor's diagonal argument: assume \( S \) is countable and list all sequences.  
Construct a new sequence differing from the \( n \)-th sequence at the \( n \)-th place.  
This sequence is not in the list — contradiction. So \( S \) is uncountable.

\begin{problembox}[2.19: Countability of Specific Sets]
Show that the following sets are countable:
\begin{itemize}
\item[(a)] Circles in the complex plane with rational radii and centers with rational coordinates.
\item[(b)] Any collection of disjoint intervals of positive length.
\end{itemize}
\end{problembox}

\textbf{Solution:}  
(a) Each circle is determined by a rational radius and two rational coordinates → set is countable.  
(b) Each disjoint interval must contain a distinct rational number → inject into \( \mathbb{Q} \), which is countable.

\begin{problembox}[2.20: Countable Support for Real Function]
Let \( f \) be a real-valued function on \( [0,1] \). Suppose there exists \( M > 0 \) such that for any finite set of points \( \{x_1, \dots, x_n\} \subset [0,1] \),  
\[
|f(x_1)| + \dots + |f(x_n)| \le M
\]  
Let \( S = \{x \in [0,1] \mid f(x) \ne 0\} \). Prove that \( S \) is countable.
\end{problembox}

\textbf{Solution:}  
Let $S = \{x \in [0,1] \mid f(x) \ne 0\}$. We want to prove that $S$ is countable.
An element $x$ is in $S$ if and only if $|f(x)| > 0$.
This is equivalent to saying that for each $x \in S$, there exists a positive integer $k$ such that $|f(x)| > 1/k$.

Let's define a collection of sets based on this idea. For each positive integer $k$, let:
\[ S_k = \left\{ x \in [0,1] \mid |f(x)| > \frac{1}{k} \right\} \]
The set $S$ is the union of all such sets:
\[ S = \bigcup_{k=1}^{\infty} S_k \]
If we can prove that each set $S_k$ is finite, then $S$ will be a countable union of finite sets, which is itself a countable set.

Let's consider a specific set $S_k$. Let $\{x_1, x_2, \dots, x_n\}$ be any finite collection of distinct points in $S_k$.
By the definition of $S_k$, we have $|f(x_i)| > 1/k$ for each $i=1, \dots, n$.
If we sum these values, we get:
\[ |f(x_1)| + |f(x_2)| + \dots + |f(x_n)| > \frac{1}{k} + \frac{1}{k} + \dots + \frac{1}{k} = \frac{n}{k} \]
The problem states that for any finite set of points, this sum is bounded by $M$:
\[ |f(x_1)| + |f(x_2)| + \dots + |f(x_n)| \le M \]
Combining these inequalities, we get:
\[ \frac{n}{k} < \sum_{i=1}^n |f(x_i)| \le M \implies \frac{n}{k} \le M \implies n \le kM \]
This result means that any finite subset of $S_k$ can have at most $kM$ elements. This implies that the set $S_k$ itself must be finite and contain at most $\lfloor kM \rfloor$ elements.

Since each $S_k$ is a finite set, their union $S = \bigcup_{k=1}^{\infty} S_k$ is a countable union of finite sets. Therefore, $S$ is a countable set.

\begin{problembox}[2.21: Fallacy in Countability of Intervals]
Find the fallacy in the following "proof" that the set of all intervals of positive length is countable:  
Let \( \{x_1, x_2, \ldots\} \) be the rationals. Every interval contains a rational \( x_n \) with minimal index \( n \).  
Assign to the interval the smallest such \( n \). This gives a function from intervals to \( \mathbb{N} \), so the set of intervals is countable.
\end{problembox}

\textbf{Solution:}  
The function \( F \) is not injective — many intervals may have the same smallest-index rational.  
So this does not establish a one-to-one correspondence between intervals and \( \mathbb{N} \).  
Hence, the proof is invalid.

\begin{problembox}[2.22: Additive Set Functions]
Let \( S \) be the collection of all subsets of a given set \( T \).  
A function \( f: S \to \mathbb{R} \) is additive if:
\[
f(A \cup B) = f(A) + f(B)
\quad \text{whenever } A \cap B = \emptyset
\]  
Prove:  
\[
f(A \cup B) = f(A) + f(B - A), \quad 
f(A \cup B) = f(A) + f(B) - f(A \cap B)
\]
\end{problembox}

\textbf{Solution:}  
Let $f: S \to \mathbb{R}$ be an additive function, meaning $f(X \cup Y) = f(X) + f(Y)$ whenever $X \cap Y = \emptyset$.

\textbf{Part 1:} Prove $f(A \cup B) = f(A) + f(B - A)$.
We can write the set $A \cup B$ as a disjoint union: $A \cup B = A \cup (B - A)$. The sets $A$ and $B-A$ (the part of $B$ not in $A$) are disjoint by definition.
Using the additivity property on this disjoint union:
\[ f(A \cup B) = f(A \cup (B-A)) = f(A) + f(B-A) \]
This proves the first identity.

\textbf{Part 2:} Prove $f(A \cup B) = f(A) + f(B) - f(A \cap B)$.
We start by decomposing $A$ and $B$ into disjoint pieces.
The set $A$ can be written as the disjoint union $A = (A - B) \cup (A \cap B)$. By additivity:
\[ f(A) = f(A - B) + f(A \cap B) \implies f(A-B) = f(A) - f(A \cap B) \]
The set $B$ can be written as the disjoint union $B = (B - A) \cup (A \cap B)$. By additivity:
\[ f(B) = f(B - A) + f(A \cap B) \implies f(B-A) = f(B) - f(A \cap B) \]
Now, we write $A \cup B$ as a union of three pairwise disjoint sets:
\[ A \cup B = (A - B) \cup (B - A) \cup (A \cap B) \]
Using the additivity property:
\[ f(A \cup B) = f(A-B) + f(B-A) + f(A \cap B) \]
Substitute the expressions we found for $f(A-B)$ and $f(B-A)$:
\begin{align*}
f(A \cup B) &= \big(f(A) - f(A \cap B)\big) + \big(f(B) - f(A \cap B)\big) + f(A \cap B) \\
&= f(A) + f(B) - f(A \cap B) - f(A \cap B) + f(A \cap B) \\
&= f(A) + f(B) - f(A \cap B)
\end{align*}
This proves the second identity.



\begin{problembox}[2.23: Solving for Total Measure from Functional Equations]

Refer to Exercise 2.22. Assume \(f\) is additive and assume also that the following relations hold for two particular subsets \(A\) and \(B\) of \(T\):
\[
f(A \cup B) = f(A') + f(B') - f(A')f(B')
\]
\[
f(A \cap B) = f(A)f(B), \quad f(A) + f(B) \ne f(T),
\]
where \(A' = T - A, B' = T - B\). Prove that these relations determine \(f(T)\), and compute the value of \(f(T)\).
\end{problembox}

\textbf{Solution:}

We are given that the function \(f\) is additive, meaning \(f(X \cup Y) = f(X) + f(Y)\) for any disjoint sets \(X\) and \(Y\). For any subset \(X \subseteq T\), its complement \(X' = T - X\) is disjoint from \(X\) and their union is \(T\). The additive property therefore implies \(f(T) = f(X) + f(X')\), which gives us:
\begin{itemize}
\item \(f(A') = f(T) - f(A)\)
\item \(f(B') = f(T) - f(B)\)
\end{itemize}
We substitute these into the first given relation:
\begin{align*}
f(A \cup B) &= \big(f(T) - f(A)\big) + \big(f(T) - f(B)\big) - \big(f(T) - f(A)\big)\big(f(T) - f(B)\big) \\
&= 2f(T) - f(A) - f(B) - \left[ f(T)^2 - f(T)f(A) - f(T)f(B) + f(A)f(B) \right] \\
&= 2f(T) - f(A) - f(B) - f(T)^2 + f(T)f(A) + f(T)f(B) - f(A)f(B)
\end{align*}
Next, we use the standard inclusion-exclusion principle for an additive function, which states \(f(A \cup B) = f(A) + f(B) - f(A \cap B)\). From the second given relation, we know \(f(A \cap B) = f(A)f(B)\). Substituting this gives:
\[
f(A \cup B) = f(A) + f(B) - f(A)f(B)
\]
Now we set the two expressions for \(f(A \cup B)\) equal to each other:
\[
f(A) + f(B) - f(A)f(B) = 2f(T) - f(A) - f(B) - f(T)^2 + f(T)f(A) + f(T)f(B) - f(A)f(B)
\]
The \( -f(A)f(B) \) terms on each side cancel. We move all remaining terms to one side to form a quadratic equation in terms of \(f(T)\):
\[
f(T)^2 - 2f(T) - f(T)f(A) - f(T)f(B) + 2f(A) + 2f(B) = 0
\]
Factoring out \(f(T)\) and constant terms:
\[
f(T)^2 - f(T)\big(2 + f(A) + f(B)\big) + 2\big(f(A) + f(B)\big) = 0
\]
This quadratic equation can be factored as:
\[
\big(f(T) - 2\big) \big(f(T) - [f(A) + f(B)]\big) = 0
\]
This implies two possible solutions for \(f(T)\):
\begin{enumerate}
\item \(f(T) = 2\)
\item \(f(T) = f(A) + f(B)\)
\end{enumerate}
Finally, we use the third given relation, \(f(A) + f(B) \ne f(T)\), to eliminate the second possibility.
Therefore, the relations uniquely determine the value of \(f(T)\). That value is 2.
\[
\boxed{f(T) = 2}
\]