\chapter{Elements of Point Set Topology}

\section{Open and Closed Sets in $\mathbb{R}^1$ and $\mathbb{R}^2$}
\begin{problembox}[3.1: Open and Closed Intervals ]
Prove that an open interval in $\mathbb{R}^1$ is an open set and that a closed interval is a closed set.
\end{problembox}

\textbf{Solution:} Let $(a,b)$ be an open interval in $\mathbb{R}^1$. To show it's open, we need to prove that every point $x \in (a,b)$ is an interior point. For any $x \in (a,b)$, let $\varepsilon = \min\{x-a, b-x\}$. Then the open ball $B(x,\varepsilon) = (x-\varepsilon, x+\varepsilon)$ is contained entirely within $(a,b)$. This shows that every point in $(a,b)$ is an interior point, so $(a,b)$ is open.

For a closed interval $[a,b]$, we need to show its complement $\mathbb{R} \setminus [a,b] = (-\infty,a) \cup (b,\infty)$ is open. Any point $x$ in this complement is either less than $a$ or greater than $b$. If $x < a$, let $\varepsilon = a-x$, then $B(x,\varepsilon) = (x-\varepsilon, x+\varepsilon) \subset (-\infty,a)$. If $x > b$, let $\varepsilon = x-b$, then $B(x,\varepsilon) \subset (b,\infty)$. This shows the complement is open, so $[a,b]$ is closed.

\begin{problembox}[3.2: Accumulation Points and Set Properties]
Determine all the accumulation points of the following sets in $\mathbb{R}^1$ and decide whether the sets are open or closed (or neither).
\begin{enumerate}[label=\textbf{(\alph*)}]
\item All integers.
\item The interval $(a, b)$.
\item All numbers of the form $1/n$,\quad $(n = 1, 2, 3, \dots)$.
\item All rational numbers.
\item All numbers of the form $2^{-n} + 5^{-m}$,\quad $(m, n = 1, 2, \dots)$.
\item All numbers of the form $(-1)^n + (1/m)$,\quad $(m, n = 1, 2, \dots)$.
\item All numbers of the form $(1/n) + (1/m)$,\quad $(m, n = 1, 2, \dots)$.
\item All numbers of the form $(-1)^n / [1 + (1/n)]$,\quad $(n = 1, 2, \dots)$
.
\end{enumerate}
\end{problembox}

\textbf{Solution:} 
(a) The set of integers has no accumulation points since each integer has a neighborhood containing no other integers. The set is closed (its complement is open) but not open.

(b) The interval $(a,b)$ has accumulation points $[a,b]$. For any $x \in [a,b]$, a sequence $\{x_n\} \subset (a,b)$ with $x_n \to x$ exists (e.g., $x_n = x + (b-a)/(n+1)$ if $x < b$, or $x_n = a + (b-a)/(n+1)$ if $x = a$). The set is open (every point is interior) but not closed (its closure is $[a,b]$).

(c) The set $\{1/n : n \in \mathbb{N}\}$ has 0 as its only accumulation point. The set is not closed, because its closure includes 0. It is also not open, as no point in the set has a neighborhood entirely contained within the set. Therefore, the set is neither open nor closed.


(d) The set of rational numbers has all real numbers as accumulation points. The set is neither open nor closed.

(e) The set $\{2^{-n} + 5^{-m} : m,n \in \mathbb{N}\}$ has accumulation points $\{2^{-n} + 5^{-m} : m,n \in \mathbb{N}\} \cup \{2^{-n} : n \in \mathbb{N}\} \cup \{5^{-m} : m \in \mathbb{N}\} \cup \{0\}$. For any $x = 2^{-k} + 5^{-l}$, take $m_n = n + l$, so $2^{-k} + 5^{-m_n} \to 2^{-k}$. Similarly, for $x = 5^{-l}$, take $n_m = m + k$, so $2^{-n_m} + 5^{-l} \to 5^{-l}$. For $x = 0$, take $n = m$, so $2^{-n} + 5^{-n} \to 0$. The set is neither open nor closed (no point is interior; closure includes accumulation points).

(g) The set $\{1/n + 1/m : m,n \in \mathbb{N}\}$ has accumulation points $\{k/n : k,n \in \mathbb{N}, k \leq n\} \cup \{0\}$. For $x = k/n$, take $m_i = i + n$, so $1/n + 1/m_i \to 1/n$; for $k/n$ with $k \geq 2$, set $m = n_i = i$, so $(k-1)/i + 1/i = k/i \to k/n$. For $x = 0$, take $n = m = i$, so $1/i + 1/i \to 0$. The set is neither open nor closed (no point is interior; closure includes accumulation points).

(h) The set $\{(-1)^n/(1+1/n) : n \in \mathbb{N}\}$ has accumulation points $\{-1, 1\}$. The set is neither open nor closed.

\begin{problembox}[3.3: Accumulation Points and Set Properties in $\mathbb{R}^2$]
The same as Exercise 3.2 for the following sets in $\mathbb{R}^2$:
\begin{enumerate}[label=\textbf{(\alph*)}]
\item All complex $z$ such that $|z| > 1$.
\item All complex $z$ such that $|z| \ge 1$.
\item All complex numbers of the form $(1/n) + (i/m)$,\quad $(m, n = 1, 2, \dots)$.
\item All points $(x, y)$ such that $x^2 - y^2 < 1$.
\item All points $(x, y)$ such that $x > 0$.
\item All points $(x, y)$ such that $x \ge 0$.
\end{enumerate}
\end{problembox}

\textbf{Solution:}
(a) The set $\{z \in \mathbb{C} : |z| > 1\}$ has accumulation points $\{z \in \mathbb{C} : |z| \geq 1\}$. The set is open but not closed.

(b) The set $\{z \in \mathbb{C} : |z| \geq 1\}$ has accumulation points $\{z \in \mathbb{C} : |z| \geq 1\}$. The set is closed but not open.

(c) The set $\{(1/n, 1/m) : m,n \in \mathbb{N}\}$ has accumulation points $\{(1/n, 0) : n \in \mathbb{N}\} \cup \{(0, 1/m) : m \in \mathbb{N}\} \cup \{(0, 0)\}$. For $(1/k, 0)$, take $(1/k, 1/m_n)$ with $m_n \to \infty$; for $(0, 1/l)$, take $(1/n_m, 1/l)$ with $n_m \to \infty$; for $(0, 0)$, take $(1/n, 1/n)$. The set is neither open nor closed (no point is interior; closure includes accumulation points).

(d) The set $\{(x,y) : x^2 - y^2 < 1\}$ has accumulation points $\{(x,y) : x^2 - y^2 \leq 1\}$. The set is open but not closed.

(e) The set $\{(x,y) : x > 0\}$ has accumulation points $\{(x,y) : x \geq 0\}$. The set is open but not closed.

(f) The set $\{(x,y) : x \geq 0\}$ has accumulation points $\{(x,y) : x \geq 0\}$. The set is closed but not open.

\begin{problembox}[3.4: Rational and Irrational Elements in Open Sets]
Prove that every nonempty open set $S$ in $\mathbb{R}^1$ contains both rational and irrational numbers.
\end{problembox}

\textbf{Solution:} Let $S$ be a nonempty open set in $\mathbb{R}^1$. Since $S$ is open, for any point $x \in S$, there exists $\varepsilon > 0$ such that the open interval $(x-\varepsilon, x+\varepsilon) \subset S$.

Since the rational numbers are dense in $\mathbb{R}$, there exists a rational number $q$ in $(x-\varepsilon, x+\varepsilon)$, and thus $q \in S$.

Similarly, since the irrational numbers are also dense in $\mathbb{R}$, there exists an irrational number $r$ in $(x-\varepsilon, x+\varepsilon)$, and thus $r \in S$.

Therefore, every nonempty open set contains both rational and irrational numbers.

\begin{problembox}[3.5: Open and Closed Sets in $\mathbb{R}^1$ and $\mathbb{R}^2$]
Prove that the only sets in $\mathbb{R}^1$ which are both open and closed are the empty set and $\mathbb{R}^1$ itself. Is a similar statement true for $\mathbb{R}^2$?
\end{problembox}

\textbf{Solution:}
This property is known as \textbf{connectedness}. A space is connected if its only subsets that are both open and closed (often called "clopen" sets) are the empty set and the space itself.

\subsubsection*{Proof for $\mathbb{R}^1$}
Let $S$ be a subset of $\mathbb{R}^1$ that is both open and closed. We'll assume, for the sake of contradiction, that $S$ is not empty and $S \neq \mathbb{R}^1$.

Since $S$ is a non-empty proper subset, its complement $S^c = \mathbb{R}^1 \setminus S$ is also non-empty. Because $S$ is closed, its complement $S^c$ is open. Because $S$ is open, its complement $S^c$ is closed. So, $S^c$ is also a non-empty clopen set.

Since both $S$ and $S^c$ are non-empty, we can choose a point $x \in S$ and a point $y \in S^c$. Without loss of generality, let's assume $x < y$.

Now, consider the set $A = S \cap [x, y]$.
\begin{itemize}
    \item The set $A$ is non-empty because $x \in A$.
    \item The set $A$ is a subset of the bounded interval $[x,y]$, so it's bounded.
    \item Since $S$ is closed and $[x, y]$ is closed, their intersection $A$ is also closed.
\end{itemize}
Because $A$ is a non-empty, closed, and bounded subset of $\mathbb{R}^1$, the least upper bound property guarantees that its supremum exists and belongs to the set. Let $s = \sup A$. Since $A$ is closed, we must have $s \in A$. This implies $s \in S$.

We know $s \le y$. Since $s \in S$ and $y \in S^c$, the points must be different, so $s < y$.

Since $s \in S$ and $S$ is an open set, there must exist an $\varepsilon > 0$ such that the open interval $(s - \varepsilon, s + \varepsilon)$ is entirely contained in $S$. We can choose $\varepsilon$ to be small enough such that $s + \varepsilon/2 < y$.
Consider the point $z = s + \varepsilon/2$.
\begin{enumerate}
    \item  $z \in (s - \varepsilon, s + \varepsilon)$, so $z \in S$.
    \item  $x \le s < z < y$, so $z \in [x, y]$.
\end{enumerate}
From these two points, $z \in S \cap [x, y]$, which means $z \in A$. However, $z > s$, which contradicts the fact that $s$ is the supremum (the least upper bound) of $A$.

This contradiction proves our initial assumption was false. Therefore, the only subsets of $\mathbb{R}^1$ that are both open and closed are $\emptyset$ and $\mathbb{R}^1$.


\subsubsection*{For $\mathbb{R}^2$}
Yes, a similar statement is true for $\mathbb{R}^2$. The only subsets of $\mathbb{R}^2$ that are both open and closed are $\emptyset$ and $\mathbb{R}^2$. This is because $\mathbb{R}^2$ is also a connected space. A common way to prove this is by showing that $\mathbb{R}^2$ is \textbf{path-connected} (any two points can be joined by a continuous path), which is a stronger condition that implies connectedness. If $\mathbb{R}^2$ were the union of two disjoint non-empty open sets, any path between a point in one set and a point in the other would have to cross the boundary, leading to a contradiction similar to the one-dimensional case.



\begin{problembox}[3.6: Closed Sets as Intersection of Open Sets]
Prove that every closed set in $\mathbb{R}^1$ is the intersection of a countable collection of open sets.
\end{problembox}

\textbf{Solution:} Let $F$ be a closed set in $\mathbb{R}^1$. For each $n \in \mathbb{N}$, define $G_n = \{x \in \mathbb{R} : d(x,F) < 1/n\}$, where $d(x,F) = \inf\{|x-y| : y \in F\}$. Each $G_n$ is open since it's the union of open intervals.

We claim that $F = \bigcap_{n=1}^{\infty} G_n$. Clearly $F \subset \bigcap_{n=1}^{\infty} G_n$ since every point in $F$ has distance 0 to $F$.

For the reverse inclusion, let $x \in \bigcap_{n=1}^{\infty} G_n$. Then $d(x,F) < 1/n$ for all $n$, which means $d(x,F) = 0$. Since $F$ is closed, this implies $x \in F$.

\begin{problembox}[3.7: Structure of Bounded Closed Sets in $\mathbb{R}^1$]
Prove that a nonempty, bounded closed set $S$ in $\mathbb{R}^1$ is either a closed interval, or that $S$ can be obtained from a closed interval by removing a countable disjoint collection of open intervals whose endpoints belong to $S$.
\end{problembox}

\textbf{Solution:} Let $S$ be a nonempty, bounded closed set in $\mathbb{R}^1$. Let $a = \inf S$ and $b = \sup S$. Since $S$ is closed, $a, b \in S$.

If $S = [a,b]$, we're done. Otherwise, the complement $[a,b] \setminus S$ is open and can be written as a countable union of disjoint open intervals $(a_i, b_i)$. Since $S$ is closed, the endpoints $a_i, b_i$ must belong to $S$.

Therefore, $S = [a,b] \setminus \bigcup_{i=1}^{\infty} (a_i, b_i)$, which is the desired representation.

\section{Open and Closed Sets in $\mathbb{R}^n$}

\begin{problembox}[3.8: Open Balls and Intervals in Rn]
Prove that open n-balls and n-dimensional open intervals are open sets in $\mathbb{R}^n$.
\end{problembox}

\textbf{Solution:} Let $B(a;r) = \{x \in \mathbb{R}^n : \|x-a\| < r\}$ be an open ball centered at $a$ with radius $r$. For any $x \in B(a;r)$, let $\varepsilon = r - \|x-a\| > 0$. Then $B(x;\varepsilon) \subset B(a;r)$ by the triangle inequality, showing $B(a;r)$ is open.

For an open interval $I = (a_1,b_1) \times \cdots \times (a_n,b_n)$, let $x = (x_1,\ldots,x_n) \in I$. For each $i$, let $\varepsilon_i = \min\{x_i - a_i, b_i - x_i\}$. Then the ball $B(x;\min\{\varepsilon_1,\ldots,\varepsilon_n\}) \subset I$, showing $I$ is open.

\begin{problembox}[3.9: Interior of a Set is Open]
Prove that the interior of a set in $\mathbb{R}^n$ is open in $\mathbb{R}^n$.
\end{problembox}

\textbf{Solution:} Let $S \subset \mathbb{R}^n$ and let $x \in \text{int } S$. By definition, there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset S$.

For any $y \in B(x;\varepsilon)$, let $\delta = \varepsilon - \|y-x\| > 0$. Then $B(y;\delta) \subset B(x;\varepsilon) \subset S$, which shows that $y \in \text{int } S$.

Therefore, $B(x;\varepsilon) \subset \text{int } S$, proving that $\text{int } S$ is open.

\begin{problembox}[3.10: Interior as Union of Open Subsets]
If $S \subseteq \mathbb{R}^n$, prove that int $S$ is the union of all open subsets of $\mathbb{R}^n$ which are contained in $S$. This is described by saying that int $S$ is the largest open subset of $S$.
\end{problembox}

\textbf{Solution:} Let $\mathcal{U}$ be the collection of all open subsets of $\mathbb{R}^n$ contained in $S$. We need to show that $\text{int } S = \bigcup_{U \in \mathcal{U}} U$.

First, if $x \in \text{int } S$, then there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset S$. Since $B(x;\varepsilon)$ is open and contained in $S$, we have $x \in B(x;\varepsilon) \in \mathcal{U}$, so $x \in \bigcup_{U \in \mathcal{U}} U$.

Conversely, if $x \in \bigcup_{U \in \mathcal{U}} U$, then $x \in U$ for some open set $U \subset S$. Since $U$ is open, there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset U \subset S$, which shows $x \in \text{int } S$.

Therefore, $\text{int } S = \bigcup_{U \in \mathcal{U}} U$, proving that the interior is the largest open subset of $S$.

\begin{problembox}[3.11: Interior of Intersection and Union]
If $S$ and $T$ are subsets of $\mathbb{R}^n$, prove that
int$(S)$ $\cap$ int$(T)$ = int$(S \cap T)$,
and int$(S)$ $\cup$ int$(T)$ $\subseteq$ int$(S \cup T)$.
\end{problembox}    

\textbf{Solution:} For the first equality, let $x \in \text{int}(S) \cap \text{int}(T)$. Then there exist $\varepsilon_1, \varepsilon_2 > 0$ such that $B(x;\varepsilon_1) \subset S$ and $B(x;\varepsilon_2) \subset T$. Let $\varepsilon = \min\{\varepsilon_1, \varepsilon_2\}$. Then $B(x;\varepsilon) \subset S \cap T$, so $x \in \text{int}(S \cap T)$.

Conversely, if $x \in \text{int}(S \cap T)$, there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset S \cap T$. This implies $B(x;\varepsilon) \subset S$ and $B(x;\varepsilon) \subset T$, so $x \in \text{int}(S) \cap \text{int}(T)$.

For the second inclusion, if $x \in \text{int}(S) \cup \text{int}(T)$, then $x \in \text{int}(S)$ or $x \in \text{int}(T)$. In either case, there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset S$ or $B(x;\varepsilon) \subset T$, which implies $B(x;\varepsilon) \subset S \cup T$. Therefore, $x \in \text{int}(S \cup T)$.

\begin{problembox}[3.12: Properties of Derived Set and Closure]
Let $S'$ denote the derived set and $\overline{S}$ the closure of a set $S$ in $\mathbb{R}^n$. Prove that:
\begin{enumerate}[label=\alph*)]
\item $S'$ is closed in $\mathbb{R}^n$; that is, $\overline{S'} \subseteq S'$.
\item If $S \subseteq T$, then $S' \subseteq T'$.
\item $S' \cup T' = (S \cup T)'$.
\item $\overline{S} = S \cup S'$.
\item $\overline{S}$ is closed in $\mathbb{R}^n$.
\item $\overline{S}$ is the intersection of all closed subsets of $\mathbb{R}^n$ containing $S$. That is, $\overline{S}$ is the smallest closed set containing $S$.
\end{enumerate}
\end{problembox}
\textbf{Solution:}\\
(a) To prove $S'$ is closed, we must show that its derived set $(S')'$ is a subset of $S'$. Let $\mathbf{x} \in (S')'$. This means every neighborhood of $\mathbf{x}$ contains a point of $S'$ other than $\mathbf{x}$. Let $B(\mathbf{x}, \varepsilon)$ be an arbitrary open ball centered at $\mathbf{x}$. By definition of $(S')'$, there is a point $\mathbf{y} \in B(\mathbf{x}, \varepsilon) \cap S'$. Since $B(\mathbf{x}, \varepsilon)$ is an open set, it is a neighborhood for $\mathbf{y}$. Because $\mathbf{y} \in S'$, $\mathbf{y}$ is an accumulation point of $S$, so this neighborhood must contain infinitely many points from $S$. Thus, the ball $B(\mathbf{x}, \varepsilon)$ contains infinitely many points from $S$. As $B(\mathbf{x}, \varepsilon)$ was an arbitrary neighborhood of $\mathbf{x}$, this shows that $\mathbf{x}$ is an accumulation point of $S$, so $\mathbf{x} \in S'$. Therefore, $(S')' \subseteq S'$, which proves that $S'$ is a closed set.

(b) Let $\mathbf{x} \in S'$. Then every neighborhood of $\mathbf{x}$ contains a point $\mathbf{y} \in S$ with $\mathbf{y} \neq \mathbf{x}$. Since $S \subseteq T$, this point $\mathbf{y}$ is also in $T$. Thus, every neighborhood of $\mathbf{x}$ contains a point $\mathbf{y} \in T$ with $\mathbf{y} \neq \mathbf{x}$. This means $\mathbf{x} \in T'$. So $S' \subseteq T'$.

(c) Using (b), since $S \subseteq S \cup T$ and $T \subseteq S \cup T$, we have $S' \subseteq (S \cup T)'$ and $T' \subseteq (S \cup T)'$. Therefore, $S' \cup T' \subseteq (S \cup T)'$. For the reverse inclusion, let $\mathbf{x} \in (S \cup T)'$. If $\mathbf{x} \notin S'$, then there is a neighborhood of $\mathbf{x}$ that contains no points of $S$ (other than possibly $\mathbf{x}$). But since $\mathbf{x} \in (S \cup T)'$, this neighborhood must contain infinitely many points from $S \cup T$. These points must therefore come from $T$. This implies $\mathbf{x} \in T'$. So, every point in $(S \cup T)'$ must be in $S'$ or $T'$. Thus, $(S \cup T)' \subseteq S' \cup T'$.

(d) The closure $\overline{S}$ consists of all points adherent to $S$. A point $\mathbf{x}$ is adherent to $S$ if every neighborhood of $\mathbf{x}$ intersects $S$.
($\overline{S} \subseteq S \cup S'$): Let $\mathbf{x} \in \overline{S}$. If $\mathbf{x} \in S$, we are done. If $\mathbf{x} \notin S$, then every neighborhood of $\mathbf{x}$ must contain a point from $S$, and that point cannot be $\mathbf{x}$. This is the definition of an accumulation point, so $\mathbf{x} \in S'$. Thus $\overline{S} \subseteq S \cup S'$.
($S \cup S' \subseteq \overline{S}$): If $\mathbf{x} \in S$, it is in $\overline{S}$ because every neighborhood contains $\mathbf{x}$. If $\mathbf{x} \in S'$, every neighborhood contains a point of $S$, so $\mathbf{x}$ is an adherent point. Thus $S' \subseteq \overline{S}$. This gives $S \cup S' \subseteq \overline{S}$.

(e) To prove $\overline{S}$ is closed, we show its derived set $(\overline{S})'$ is a subset of $\overline{S}$. From (d), $\overline{S} = S \cup S'$. Using (c), we get $(\overline{S})' = (S \cup S')' = S' \cup (S')'$. From (a), $S'$ is closed, which means $(S')' \subseteq S'$. Therefore, $(\overline{S})' \subseteq S' \cup S' = S'$. Since $S' \subseteq S \cup S' = \overline{S}$, we have $(\overline{S})' \subseteq \overline{S}$. This proves that $\overline{S}$ is closed.

(f) Let $\mathcal{C}$ be the collection of all closed sets containing $S$. Let $C_{min} = \bigcap_{F \in \mathcal{C}} F$.
($\overline{S} \subseteq C_{min}$): Let $F$ be any set in $\mathcal{C}$. Then $F$ is closed and $S \subseteq F$. The closure of a set is the smallest closed set containing it, so we must have $\overline{S} \subseteq F$. Since this holds for all $F \in \mathcal{C}$, we have $\overline{S} \subseteq \bigcap_{F \in \mathcal{C}} F = C_{min}$.
($C_{min} \subseteq \overline{S}$): By part (e), $\overline{S}$ is a closed set. It also contains $S$. Therefore, $\overline{S}$ is one of the sets in the collection $\mathcal{C}$. The intersection of all sets in $\mathcal{C}$ must be a subset of any particular member, so $C_{min} \subseteq \overline{S}$.
Thus, $\overline{S} = C_{min}$.

\begin{problembox}[3.13: Closure under Intersection of Sets]

Let $S$ and $T$ be subsets of $\mathbb{R}^k$. Prove that $\overline{S \cup T} = \overline{S} \cup \overline{T}$ and that $\overline{S \cap T} \subseteq \overline{S} \cap \overline{T}$ if $S$ is open.

\textit{NOTE.} The statements in Exercises 3.9 through 3.13 are true in any metric space.
\end{problembox}

\textbf{Solution:}
\subsubsection*{Closure of a Union: $\overline{S \cup T} = \overline{S} \cup \overline{T}$}
A concise way to prove this is by using the results from Exercise 3.12:
\begin{align*}
\overline{S \cup T} &= (S \cup T) \cup (S \cup T)' && (\text{by 3.12d}) \\
&= (S \cup T) \cup (S' \cup T') && (\text{by 3.12c}) \\
&= (S \cup S') \cup (T \cup T') && (\text{by associativity and commutativity of } \cup) \\
&= \overline{S} \cup \overline{T} && (\text{by 3.12d})
\end{align*}

\subsubsection*{Closure of an Intersection: $\overline{S \cap T} \subseteq \overline{S} \cap \overline{T}$}
Let $\mathbf{x} \in \overline{S \cap T}$. By the definition of closure, this means every neighborhood of $\mathbf{x}$ contains at least one point from $S \cap T$.
Let $U$ be any neighborhood of $\mathbf{x}$. Then $U \cap (S \cap T) \neq \emptyset$.
Since $S \cap T \subseteq S$, it follows that $U \cap S \neq \emptyset$. As $U$ was an arbitrary neighborhood of $\mathbf{x}$, this implies that $\mathbf{x}$ is an adherent point of $S$, so $\mathbf{x} \in \overline{S}$.
Similarly, since $S \cap T \subseteq T$, it follows that $U \cap T \neq \emptyset$. This implies that $\mathbf{x}$ is an adherent point of $T$, so $\mathbf{x} \in \overline{T}$.
Since $\mathbf{x} \in \overline{S}$ and $\mathbf{x} \in \overline{T}$, we have $\mathbf{x} \in \overline{S} \cap \overline{T}$.
Therefore, $\overline{S \cap T} \subseteq \overline{S} \cap \overline{T}$.

\textit{Note: The inclusion can be strict. For example, in $\mathbb{R}^1$, let $S = \mathbb{Q}$ and $T = \mathbb{R} \setminus \mathbb{Q}$. Then $S \cap T = \emptyset$, so $\overline{S \cap T} = \emptyset$. However, $\overline{S} = \mathbb{R}$ and $\overline{T} = \mathbb{R}$, so $\overline{S} \cap \overline{T} = \mathbb{R}$.}

\begin{problembox}[3.14: Properties of Convex Sets]
A set \( S \) in \( \mathbb{R}^n \) is called convex if, for every pair of points \( x \) and \( y \) in \( S \) and every real \( \theta \) satisfying \( 0 < \theta < 1 \), we have \( \theta x + (1 - \theta)y \in S \). Interpret this statement geometrically (in \( \mathbb{R}^2 \) and \( \mathbb{R}^3 \)) and prove that:
\begin{enumerate}[label=\alph*)]
\item Every \( n \)-ball in \( \mathbb{R}^n \) is convex.
\item Every \( n \)-dimensional open interval is convex.
\item The interior of a convex set is convex.
\item The closure of a convex set is convex.
\end{enumerate}
\end{problembox}

\textbf{Solution:} Geometrically, a set is convex if the line segment joining any two points in the set lies entirely within the set.

(a) Let $B(a;r)$ be an $n$-ball and $x, y \in B(a;r)$. For $0 < \theta < 1$, let $z = \theta x + (1-\theta)y$. Then $\|z-a\| = \|\theta(x-a) + (1-\theta)(y-a)\| \leq \theta\|x-a\| + (1-\theta)\|y-a\| < \theta r + (1-\theta)r = r$, so $z \in B(a;r)$.

(b) Let $I = (a_1,b_1) \times \cdots \times (a_n,b_n)$ be an open interval and $x, y \in I$. For $0 < \theta < 1$, let $z = \theta x + (1-\theta)y$. For each $i$, we have $a_i < x_i, y_i < b_i$, so $a_i < \theta x_i + (1-\theta)y_i < b_i$. Therefore, $z \in I$.

(c) Let $S$ be convex and $x, y \in \text{int } S$. There exist $\varepsilon_1, \varepsilon_2 > 0$ such that $B(x;\varepsilon_1) \subset S$ and $B(y;\varepsilon_2) \subset S$. Let $\varepsilon = \min\{\varepsilon_1, \varepsilon_2\}$. For $0 < \theta < 1$, let $z = \theta x + (1-\theta)y$. If $w \in B(z;\varepsilon)$, then $\|w-z\| < \varepsilon$. Let $u = w - z + x$ and $v = w - z + y$. Then $\|u-x\| = \|v-y\| = \|w-z\| < \varepsilon$, so $u, v \in S$. Since $S$ is convex, $w = \theta u + (1-\theta)v \in S$. Therefore, $B(z;\varepsilon) \subset S$, so $z \in \text{int } S$.

(d) Let $S$ be convex and $x, y \in \overline{S}$. There exist sequences $\{x_n\}, \{y_n\} \subset S$ converging to $x, y$ respectively. For $0 < \theta < 1$, let $z = \theta x + (1-\theta)y$ and $z_n = \theta x_n + (1-\theta)y_n$. Since $S$ is convex, $z_n \in S$ for all $n$. Since $z_n \to z$, we have $z \in \overline{S}$.

\begin{problembox}[3.15: Accumulation Points of Intersections and Unions]            
Let $\mathcal{F}$ be a collection of sets in $\mathbb{R}^k$, and let $S = \bigcup_{A \in \mathcal{F}} A$ and $T = \bigcap_{A \in \mathcal{F}} A$. For each of the following statements, either give a proof or exhibit a counterexample:
\begin{enumerate}[label=\alph*)]
\item If $\mathbf{x}$ is an accumulation point of $T$, then $\mathbf{x}$ is an accumulation point of each set $A$ in $\mathcal{F}$.
\item If $\mathbf{x}$ is an accumulation point of $S$, then $\mathbf{x}$ is an accumulation point of at least one set $A$ in $\mathcal{F}$.
\end{enumerate}
\end{problembox}

\textbf{Solution:}\\
(a) This statement is \textbf{true}.\\
\textbf{Proof:} Let $\mathbf{x}$ be an accumulation point of $T$. This means that for any $\varepsilon > 0$, the ball $B(\mathbf{x}; \varepsilon)$ contains a point $\mathbf{y} \in T$ such that $\mathbf{y} \neq \mathbf{x}$. By definition, $T = \bigcap_{A \in \mathcal{F}} A$. So, if $\mathbf{y} \in T$, then $\mathbf{y} \in A$ for every set $A$ in the collection $\mathcal{F}$. Therefore, for any $\varepsilon > 0$, the ball $B(\mathbf{x}; \varepsilon)$ contains a point $\mathbf{y} \in A$ (for every $A \in \mathcal{F}$) with $\mathbf{y} \neq \mathbf{x}$. This is precisely the definition of $\mathbf{x}$ being an accumulation point of $A$. Thus, $\mathbf{x}$ is an accumulation point of each set $A \in \mathcal{F}$.

(b) This statement is \textbf{false} for an infinite collection $\mathcal{F}$.\\
\textbf{Counterexample:} Let the collection of sets in $\mathbb{R}^1$ be $\mathcal{F} = \{A_n : n \in \mathbb{N}\}$ where each set $A_n$ is a singleton: $A_n = \{1/n\}$.
The union is the set $S = \bigcup_{n=1}^{\infty} A_n = \{1, 1/2, 1/3, \dots\}$.
The set $S$ has exactly one accumulation point: $0$, since the sequence of points converges to $0$.
However, none of the individual sets $A_n$ have any accumulation points, as they each contain only a single isolated point.
Thus, $0$ is an accumulation point of $S$, but not of any set $A_n$ in the collection $\mathcal{F}$.

\textit{(Note: The statement is true if the collection $\mathcal{F}$ is finite. If $\mathbf{x}$ is an accumulation point of a finite union $S = A_1 \cup \dots \cup A_m$, then any neighborhood of $\mathbf{x}$ contains infinitely many points from $S$. By the pigeonhole principle, at least one of the sets $A_i$ must contribute infinitely many of these points, making $\mathbf{x}$ an accumulation point of that $A_i$.)}


\begin{problembox}[3.16: Rationals  Not a Countable Intersection of Open Sets]
Prove that the set \( S \) of rational numbers in the interval \( (0, 1) \) cannot be expressed as the intersection of a countable collection of open sets. 

\textit{Hint.} Write \( S = \{x_1, x_2, \ldots\} \), assume \( S = \bigcap_{k=1}^{\infty} S_k \), where each \( S_k \) is open, and construct a sequence \( (Q_n) \) of closed intervals such that \( Q_{n+1} \subseteq Q_n \subseteq S_n \) and such that \( x_n \notin Q_n \). Then use the Cantor intersection theorem to obtain a contradiction.
\end{problembox}

\textbf{Solution:} Suppose for contradiction that $S = \bigcap_{k=1}^{\infty} S_k$ where each $S_k$ is open. Let $S = \{x_1, x_2, \ldots\}$ be an enumeration of the rationals in $(0,1)$.

For each $n$, since $S_n$ is open and contains all rationals in $(0,1)$, we can find a closed interval $Q_n \subset S_n$ such that $x_n \notin Q_n$. We can arrange that $Q_{n+1} \subseteq Q_n$ by taking $Q_{n+1} = Q_n \cap I_{n+1}$ where $I_{n+1}$ is a closed interval in $S_{n+1}$ that doesn't contain $x_{n+1}$.

By the Cantor intersection theorem, $\bigcap_{n=1}^{\infty} Q_n$ is nonempty. Let $x \in \bigcap_{n=1}^{\infty} Q_n$. Then $x \in \bigcap_{k=1}^{\infty} S_k = S$, so $x$ is rational. But $x \neq x_n$ for any $n$ since $x_n \notin Q_n$ for each $n$. This contradicts the fact that $S$ contains all rationals in $(0,1)$.

\section{Covering Theorems in $\mathbb{R}^n$}

\begin{problembox}[3.17: Countability of Isolated Points]
If \( S \subseteq \mathbb{R}^n \), prove that the collection of isolated points of \( S \) is countable.
\end{problembox}

\textbf{Solution:}\\
Let $I$ be the set of isolated points of $S$. By definition, for each point $\mathbf{x} \in I$, there exists a radius $\varepsilon_{\mathbf{x}} > 0$ such that the open ball $B(\mathbf{x}; \varepsilon_{\mathbf{x}})$ contains no other point of $S$; that is, $B(\mathbf{x}; \varepsilon_{\mathbf{x}}) \cap S = \{\mathbf{x}\}$.

Consider the collection of smaller open balls $\mathcal{C} = \{ B(\mathbf{x}; \varepsilon_{\mathbf{x}}/2) : \mathbf{x} \in I \}$. We claim these balls are pairwise disjoint.
To prove this, let $\mathbf{x}_1$ and $\mathbf{x}_2$ be two distinct points in $I$. Suppose their corresponding balls in $\mathcal{C}$ have a point $\mathbf{y}$ in common. Then $d(\mathbf{x}_1, \mathbf{y}) < \varepsilon_{\mathbf{x}_1}/2$ and $d(\mathbf{x}_2, \mathbf{y}) < \varepsilon_{\mathbf{x}_2}/2$.
By the triangle inequality:
$$d(\mathbf{x}_1, \mathbf{x}_2) \le d(\mathbf{x}_1, \mathbf{y}) + d(\mathbf{y}, \mathbf{x}_2) < \frac{\varepsilon_{\mathbf{x}_1}}{2} + \frac{\varepsilon_{\mathbf{x}_2}}{2}$$
Assuming, without loss of generality, that $\varepsilon_{\mathbf{x}_1} \le \varepsilon_{\mathbf{x}_2}$, we get $d(\mathbf{x}_1, \mathbf{x}_2) < \varepsilon_{\mathbf{x}_2}/2 + \varepsilon_{\mathbf{x}_2}/2 = \varepsilon_{\mathbf{x}_2}$.
This implies that $\mathbf{x}_1 \in B(\mathbf{x}_2; \varepsilon_{\mathbf{x}_2})$. But $\mathbf{x}_1 \in S$ and $\mathbf{x}_1 \neq \mathbf{x}_2$. This contradicts the fact that $B(\mathbf{x}_2; \varepsilon_{\mathbf{x}_2})$ contains only one point from $S$, namely $\mathbf{x}_2$.
Therefore, the balls in the collection $\mathcal{C}$ must be pairwise disjoint.

Now we use the fact that $\mathbb{R}^n$ is \textbf{separable}, meaning it contains a countable dense subset, such as $\mathbb{Q}^n$ (the set of points with rational coordinates).
Since each ball in $\mathcal{C}$ is a non-empty open set, each must contain at least one point from the dense set $\mathbb{Q}^n$. Because the balls in $\mathcal{C}$ are disjoint, each ball must contain a \textit{different} rational point.
This allows us to define an injective (one-to-one) function from the set of isolated points $I$ to the countable set $\mathbb{Q}^n$ (by mapping each $\mathbf{x} \in I$ to a rational point in $B(\mathbf{x}; \varepsilon_{\mathbf{x}}/2)$). A set that can be mapped injectively into a countable set must itself be countable.
Thus, the set of isolated points $I$ is countable.

\begin{problembox}[3.18: Countable Covering of the First Quadrant]
Prove that the set of open disks in the \(xy\)-plane with center at \( (x, x) \) and radius \( x > 0 \), where \( x \) is rational, is a countable covering of the set \( \{(x, y) : x > 0, y > 0\} \).
\end{problembox}

\textbf{Solution:}\\
Let $\mathcal{F}$ be the collection of open disks $B((q,q); q)$ where $q \in \mathbb{Q}$ and $q > 0$. Since $\mathbb{Q}$ is countable, the collection $\mathcal{F}$ is countable. We need to show that $\mathcal{F}$ covers the first quadrant $S = \{(x, y) : x > 0, y > 0\}$.

Let $(x, y)$ be an arbitrary point in $S$. We need to find a rational number $q > 0$ such that the disk $B((q,q); q)$ contains $(x, y)$. The condition for this is:
$$\sqrt{(x-q)^2 + (y-q)^2} < q$$
Since both sides are positive, we can square the inequality:
$$(x-q)^2 + (y-q)^2 < q^2$$
$$x^2 - 2xq + q^2 + y^2 - 2yq + q^2 < q^2$$
$$q^2 - 2(x+y)q + (x^2+y^2) < 0$$
Let $f(q) = q^2 - 2(x+y)q + (x^2+y^2)$. We are looking for a rational $q > 0$ that makes this quadratic expression negative. The graph of $z=f(q)$ is an upward-opening parabola. It will be negative between its roots. The roots are found using the quadratic formula:
$$q = \frac{2(x+y) \pm \sqrt{4(x+y)^2 - 4(x^2+y^2)}}{2} = (x+y) \pm \sqrt{(x+y)^2 - (x^2+y^2)}$$
$$q = (x+y) \pm \sqrt{2xy}$$
Let the roots be $q_1 = (x+y) - \sqrt{2xy}$ and $q_2 = (x+y) + \sqrt{2xy}$. Since $x,y > 0$, the term $\sqrt{2xy}$ is real and positive, so $q_1 < q_2$. The interval $(q_1, q_2)$ is non-empty.
Since the rational numbers are dense in $\mathbb{R}$, we can always find a rational number $q$ in this interval: $q_1 < q < q_2$. For any such $q$, the inequality $f(q) < 0$ holds.

We must also ensure that we can choose $q$ to be positive. The product of the roots is $q_1 q_2 = x^2+y^2 > 0$. Since $q_2 = (x+y) + \sqrt{2xy}$ is clearly positive, the other root $q_1$ must also be positive.
Since the interval $(q_1, q_2)$ consists of positive numbers and contains a rational number, we can always find a suitable rational $q > 0$.
Thus, for any point $(x,y)$ in the first quadrant, we can find a disk in $\mathcal{F}$ that contains it. The countable collection $\mathcal{F}$ therefore covers the first quadrant.


\begin{problembox}[3.19: Non-Finite Subcover of \(0,1\)]
The collection \( \mathcal{F} \) of open intervals of the form \( (1/n, 2/n) \), where \( n = 2, 3, \ldots \), is an open covering of the open interval \( (0, 1) \). Prove (without using Theorem 3.31) that no finite subcollection of \( \mathcal{F} \) covers \( (0, 1) \).
\end{problembox}

\textbf{Solution:} Let $\mathcal{G} = \{(1/n_1, 2/n_1), \ldots, (1/n_k, 2/n_k)\}$ be a finite subcollection of $\mathcal{F}$. Let $N = \max\{n_1, \ldots, n_k\}$.

Then the largest interval in $\mathcal{G}$ is $(1/N, 2/N)$. For any $x \in (0, 1/N)$, we have $x < 1/N < 2/N$, so $x$ is not covered by any interval in $\mathcal{G}$.

Therefore, $\mathcal{G}$ does not cover $(0,1)$, proving that no finite subcollection of $\mathcal{F}$ covers $(0,1)$.

\begin{problembox}[3.20: Closed but Not Bounded Set with Infinite Covering]
Give an example of a set \( S \) which is closed but not bounded and exhibit a countable open covering \( \mathcal{F} \) such that no finite subset of \( \mathcal{F} \) covers \( S \).
\end{problembox}

\textbf{Solution:} Let $S = \mathbb{Z}$ (the set of integers). This set is closed but not bounded.

Let $\mathcal{F} = \{(n-1/2, n+1/2) : n \in \mathbb{Z}\}$. This is a countable open covering of $\mathbb{Z}$ since each integer $n$ is contained in the interval $(n-1/2, n+1/2)$.

However, no finite subcollection of $\mathcal{F}$ covers $\mathbb{Z}$. If $\mathcal{G} = \{(n_1-1/2, n_1+1/2), \ldots, (n_k-1/2, n_k+1/2)\}$ is a finite subcollection, then $\mathcal{G}$ can only cover finitely many integers, but $\mathbb{Z}$ is infinite.

Therefore, $\mathcal{F}$ is a countable open covering of $S$ with no finite subcover.

\begin{problembox}[3.21: Countability via Local Countability]
Given a set \( S \) in \( \mathbb{R}^n \) with the property that for every \( x \) in \( S \) there is an \( n \)-ball \( B(x) \) such that \( B(x) \cap S \) is countable. Prove that \( S \) is countable.
\end{problembox}

\textbf{Solution:}\\
For each point $\mathbf{x} \in S$, we are given that there exists an open ball $B_\mathbf{x}$ centered at $\mathbf{x}$ such that the set $B_\mathbf{x} \cap S$ is countable.

The collection of all such balls, $\mathcal{C} = \{B_\mathbf{x} : \mathbf{x} \in S\}$, forms an open covering of the set $S$ (since each $\mathbf{x} \in S$ is in its own ball $B_\mathbf{x}$).

The space $\mathbb{R}^n$ is a \textbf{separable} metric space because it contains a countable dense subset, $\mathbb{Q}^n$. A key theorem in topology states that every separable metric space has the \textbf{Lindelöf property}. This property guarantees that any open covering of a set in that space has a countable subcovering.

Applying the Lindelöf property to our open cover $\mathcal{C}$ of $S$, we can extract a countable subcollection, say $\mathcal{C}' = \{B_{\mathbf{x}_k} : k \in \mathbb{N}\}$, that still covers $S$. This means:
$$S \subseteq \bigcup_{k=1}^{\infty} B_{\mathbf{x}_k}$$
From this, we can express the set $S$ as:
$$S = S \cap \left( \bigcup_{k=1}^{\infty} B_{\mathbf{x}_k} \right) = \bigcup_{k=1}^{\infty} (S \cap B_{\mathbf{x}_k})$$
By the initial hypothesis, each set in this union, $(S \cap B_{\mathbf{x}_k})$, is countable.
Therefore, $S$ is a countable union of countable sets. A fundamental result of set theory states that a countable union of countable sets is itself countable. Thus, we conclude that the set $S$ must be countable.


\begin{problembox}[3.22: Countability of Disjoint Open Sets]
Prove that a collection of disjoint open sets in \( \mathbb{R}^n \) is necessarily countable. Give an example of a collection of disjoint closed sets which is not countable.
\end{problembox}

\textbf{Solution:} Let $\mathcal{F}$ be a collection of disjoint open sets in $\mathbb{R}^n$. Since $\mathbb{R}^n$ is separable, there exists a countable dense subset $D$.

For each open set $U \in \mathcal{F}$, there exists a point $d \in D$ such that $d \in U$. Since the sets in $\mathcal{F}$ are disjoint, each point $d \in D$ can belong to at most one set in $\mathcal{F}$.

Therefore, the number of sets in $\mathcal{F}$ is at most the number of points in $D$, which is countable.

For an example of uncountably many disjoint closed sets, let $\mathcal{G} = \{\{x\} : x \in \mathbb{R}\}$. Each singleton $\{x\}$ is closed, the sets are disjoint, and there are uncountably many real numbers.

\begin{problembox}[3.23: Existence of Condensation Points]
Assume that \( S \subseteq \mathbb{R}^n \). A point \( x \) in \( \mathbb{R}^n \) is said to be a condensation point of \( S \) if every \( n \)-ball \( B(x) \) has the property that \( B(x) \cap S \) is not countable. Prove that if \( S \) is not countable, then there exists a point \( x \) in \( S \) such that \( x \) is a condensation point of \( S \).
\end{problembox}

\textbf{Solution:} Suppose for contradiction that no point in $S$ is a condensation point of $S$. Then for every $x \in S$, there exists an $n$-ball $B_x$ centered at $x$ such that $B_x \cap S$ is countable.

By Exercise 3.21, this implies that $S$ is countable, which contradicts the hypothesis that $S$ is not countable.

Therefore, there must exist at least one point $x \in S$ that is a condensation point of $S$.

\begin{problembox}[3.24: Properties of Condensation Points]
Assume that \( S \subseteq \mathbb{R}^n \) and that \( S \) is not countable. Let \( T \) denote the set of condensation points of \( S \). Prove that:
\begin{enumerate}[label=\alph*)]
\item \( S - T \) is countable,
\item \( S \cap T \) is not countable,
\item \( T \) is a closed set,
\item \( T \) contains no isolated points.
\end{enumerate}
Note that Exercise 3.23 is a special case of (b).
\end{problembox}

\textbf{Solution:} 
(a) For each $x \in S - T$, there exists an $n$-ball $B_x$ centered at $x$ such that $B_x \cap S$ is countable. By Exercise 3.21, $S - T$ is countable.

(b) Since $S$ is not countable and $S - T$ is countable, $S \cap T$ must be uncountable.

(c) Let $x \in \overline{T}$. Then every neighborhood of $x$ contains a point of $T$. Let $B$ be any $n$-ball centered at $x$. There exists $y \in T \cap B$. Since $y$ is a condensation point, $B(y;r) \cap S$ is uncountable for any $r > 0$. Choose $r$ small enough so that $B(y;r) \subset B$. Then $B \cap S$ contains the uncountable set $B(y;r) \cap S$, so $x$ is a condensation point. Therefore, $T$ is closed.

(d) Let $x \in T$. For any $\varepsilon > 0$, $B(x;\varepsilon) \cap S$ is uncountable. Since $S - T$ is countable, $B(x;\varepsilon) \cap T$ must be uncountable. Therefore, $x$ is not isolated in $T$.

\begin{problembox}[3.25: Cantor-Bendixon Theorem]
A set in \( \mathbb{R}^n \) is called perfect if \( S = S' \), that is, if \( S \) is a closed set which contains no isolated points. Prove that every uncountable closed set \( F \) in \( \mathbb{R}^n \) can be expressed in the form \( F = A \cup B \), where \( A \) is perfect and \( B \) is countable (Cantor-Bendixon theorem).

\textit{Hint.} Use Exercise 3.24.
\end{problembox}

\textbf{Solution:} Let $F$ be an uncountable closed set in $\mathbb{R}^n$. Let $T$ be the set of condensation points of $F$. By Exercise 3.24, $T$ is closed and $F - T$ is countable.

Let $A = T$ and $B = F - T$. Then $F = A \cup B$ where $B$ is countable.

We need to show that $A$ is perfect. Since $T$ is closed by Exercise 3.24(c), $A$ is closed. By Exercise 3.24(d), $T$ contains no isolated points, so $A$ contains no isolated points.

Therefore, $A$ is perfect, and we have the desired decomposition $F = A \cup B$.

\section{Metric Spaces}

\begin{problembox}[3.26: Open and Closed Sets in Metric Spaces]
In any metric space \((M, d)\), prove that the empty set \( \emptyset \) and the whole space \( M \) are both open and closed.
\end{problembox}

\textbf{Solution:} The empty set $\emptyset$ is open because the condition "for every point in $\emptyset$, there exists a neighborhood contained in $\emptyset$" is vacuously true (there are no points to check).

The empty set $\emptyset$ is closed because its complement $M$ is open.

The whole space $M$ is open because for any point $x \in M$ and any $\varepsilon > 0$, the ball $B(x;\varepsilon) \subset M$.

The whole space $M$ is closed because its complement $\emptyset$ is open.

\begin{problembox}[3.27: Metric Balls in Different Metrics]
Consider the following two metrics in \( \mathbb{R}^n \):
\[d_1(x, y) = \max_{1 \leq i \leq n} |x_i - y_i|, \quad d_2(x, y) = \sum_{i=1}^n |x_i - y_i|.\]

In each of the following metric spaces prove that the ball \( B(a; r) \) has the geometric appearance indicated:
\begin{enumerate}[label=\alph*)]
\item In \( (\mathbb{R}^2, d_1) \), a square with sides parallel to the coordinate axes.
\item In \( (\mathbb{R}^2, d_2) \), a square with diagonals parallel to the axes.
\item A cube in \( (\mathbb{R}^3, d_1) \).
\item An octahedron in \( (\mathbb{R}^3, d_2) \).
\end{enumerate}
\end{problembox}

\textbf{Solution:} 
(a) In $(\mathbb{R}^2, d_1)$, the ball $B(a;r) = \{(x,y) : \max\{|x-a_1|, |y-a_2|\} < r\}$. This means $|x-a_1| < r$ and $|y-a_2| < r$, which defines a square with center $(a_1,a_2)$ and sides of length $2r$ parallel to the coordinate axes.

(b) In $(\mathbb{R}^2, d_2)$, the ball $B(a;r) = \{(x,y) : |x-a_1| + |y-a_2| < r\}$. This defines a diamond-shaped region (square rotated 45 degrees) with diagonals parallel to the axes.

(c) In $(\mathbb{R}^3, d_1)$, the ball $B(a;r) = \{(x,y,z) : \max\{|x-a_1|, |y-a_2|, |z-a_3|\} < r\}$. This defines a cube with center $(a_1,a_2,a_3)$ and sides of length $2r$ parallel to the coordinate axes.

(d) In $(\mathbb{R}^3, d_2)$, the ball $B(a;r) = \{(x,y,z) : |x-a_1| + |y-a_2| + |z-a_3| < r\}$. This defines an octahedron with center $(a_1,a_2,a_3)$.

\begin{problembox}[3.28: Metric Inequalities]
Let \( d_1 \) and \( d_2 \) be the metrics of Exercise 3.27 and let \( \|x - y\| \) denote the usual Euclidean metric. Prove the following inequalities for all \( x \) and \( y \) in \( \mathbb{R}^n \):
\[d_1(x, y) \leq \|x - y\| \leq d_2(x, y) \quad \text{and} \quad d_2(x, y) \leq \sqrt{n} \|x - y\| \leq n\,d_1(x, y).\]
\end{problembox}

\textbf{Solution:} Let $x, y \in \mathbb{R}^n$. Let $a_i = |x_i - y_i|$ for $1 \leq i \leq n$. Then:

(1) \( d_1(x, y) = \max_i a_i \leq \sqrt{\sum a_i^2} = \|x - y\| \) (since each $a_i^2 \leq \sum a_i^2$)

(2) \( \|x - y\| = \sqrt{\sum a_i^2} \leq \sum a_i = d_2(x,y) \) (by the inequality \( \sqrt{a_1^2 + \dots + a_n^2} \leq a_1 + \dots + a_n \))

(3) By the Cauchy-Schwarz inequality:
\[
\left(\sum a_i\right)^2 \leq n \sum a_i^2 \Rightarrow d_2(x,y) \leq \sqrt{n} \|x - y\|
\]

(4) Also, since \( \|x - y\| = \sqrt{\sum a_i^2} \leq \sqrt{n \cdot \max_i a_i^2} = \sqrt{n} \cdot \max a_i = \sqrt{n} d_1(x, y) \), it follows that:

\[
\sqrt{n} \|x - y\| \leq n d_1(x, y)
\]

Hence, all inequalities hold:

\[
d_1(x,y) \leq \|x - y\| \leq d_2(x,y), \quad d_2(x, y) \leq \sqrt{n} \|x - y\| \leq n\,d_1(x, y)
\]


\begin{problembox}[3.29: Bounded Metric]
If \( (M, d) \) is a metric space, define
\[d'(x, y) = \frac{d(x, y)}{1 + d(x, y)}.\]
Prove that \( d' \) is also a metric for \( M \). Note that \( 0 \leq d'(x, y) < 1 \) for all \( x, y \) in \( M \).
\end{problembox}

\textbf{Solution:} We need to verify the three properties of a metric:

(1) $d'(x,y) \geq 0$ since $d(x,y) \geq 0$ and $1 + d(x,y) > 0$.

(2) $d'(x,y) = 0$ if and only if $d(x,y) = 0$, which occurs if and only if $x = y$.

(3) $d'(x,y) = d'(y,x)$ since $d(x,y) = d(y,x)$.

(4) For the triangle inequality, let $f(t) = \frac{t}{1+t}$. Then $f'(t) = \frac{1}{(1+t)^2} > 0$, so $f$ is increasing. Therefore, $d'(x,z) = f(d(x,z)) \leq f(d(x,y) + d(y,z)) = \frac{d(x,y) + d(y,z)}{1 + d(x,y) + d(y,z)} \leq \frac{d(x,y)}{1 + d(x,y)} + \frac{d(y,z)}{1 + d(y,z)} = d'(x,y) + d'(y,z)$.

The last inequality follows from the fact that $\frac{a+b}{1+a+b} \leq \frac{a}{1+a} + \frac{b}{1+b}$ for $a,b \geq 0$.

\begin{problembox}[3.30: Finite Sets in Metric Spaces]
Prove that every finite subset of a metric space is closed.
\end{problembox}

\textbf{Solution:} Let $S = \{x_1, x_2, \ldots, x_n\}$ be a finite subset of a metric space $(M,d)$. We need to show that the complement $M \setminus S$ is open.

Let $x \in M \setminus S$. Let $\varepsilon = \min\{d(x,x_i) : i = 1,2,\ldots,n\}$. Since $x \notin S$, we have $\varepsilon > 0$.

Then $B(x;\varepsilon) \cap S = \emptyset$, so $B(x;\varepsilon) \subset M \setminus S$. This shows that every point in $M \setminus S$ is an interior point, so $M \setminus S$ is open.

Therefore, $S$ is closed.

\begin{problembox}[3.31: Closed Balls in Metric Spaces]
In a metric space \((M, d)\) the closed ball of radius \( r > 0 \) about a point \( a \) in \( M \) is the set \( \overline{B}(a; r) = \{x : d(x, a) \leq r\} \).
\begin{enumerate}[label=\alph*)]
\item Prove that \( \overline{B}(a; r) \) is a closed set.
\item Give an example of a metric space in which \( \overline{B}(a; r) \) is not the closure of the open ball \( B(a; r) \).
\end{enumerate}
\end{problembox}

\textbf{Solution:} 
(a) Let $x \in M \setminus \overline{B}(a;r)$. Then $d(x,a) > r$. Let $\varepsilon = d(x,a) - r > 0$. For any $y \in B(x;\varepsilon)$, we have $d(y,a) \geq d(x,a) - d(x,y) > d(x,a) - \varepsilon = r$. Therefore, $B(x;\varepsilon) \subset M \setminus \overline{B}(a;r)$, showing that $M \setminus \overline{B}(a;r)$ is open. Hence, $\overline{B}(a;r)$ is closed.

(b) Consider the discrete metric space $(M,d)$ where $d(x,y) = 1$ if $x \neq y$ and $d(x,y) = 0$ if $x = y$. Let $a \in M$ and $r = 1$. Then $B(a;1) = \{a\}$ and $\overline{B}(a;1) = M$. The closure of $B(a;1)$ is $\{a\}$, which is not equal to $\overline{B}(a;1) = M$.

\begin{problembox}[3.32: Transitivity of Density]
In a metric space \( M \), if subsets satisfy \( A \subseteq S \subseteq \overline{A} \), where \(\overline{A}\) is the closure of \( A \), then \( A \) is said to be dense in \( S \). For example, the set \( \mathbb{Q} \) of rational numbers is dense in \( \mathbb{R} \). If \( A \) is dense in \( S \) and if \( S \) is dense in \( T \), prove that \( A \) is dense in \( T \).
\end{problembox}

\textbf{Solution:} We need to show that $A \subseteq T \subseteq \overline{A}$.

Since $A \subseteq S \subseteq T$, we have $A \subseteq T$.

Since $S$ is dense in $T$, we have $T \subseteq \overline{S}$. Since $A$ is dense in $S$, we have $S \subseteq \overline{A}$. Therefore, $\overline{S} \subseteq \overline{\overline{A}} = \overline{A}$.

Combining these, we get $T \subseteq \overline{S} \subseteq \overline{A}$, so $T \subseteq \overline{A}$.

Therefore, $A \subseteq T \subseteq \overline{A}$, showing that $A$ is dense in $T$.

\begin{problembox}[3.33: Separability of Euclidean Spaces]
A metric space \( M \) is said to be separable if there is a countable subset \( A \) which is dense in \( M \). For example, \( \mathbb{R} \) is separable because the set \( \mathbb{Q} \) of rational numbers is a countable dense subset. Prove that every Euclidean space \( \mathbb{R}^k \) is separable.
\end{problembox}

\textbf{Solution:} Let $A$ be the set of all points in $\mathbb{R}^k$ with rational coordinates. That is, $A = \{(q_1, q_2, \ldots, q_k) : q_i \in \mathbb{Q} \text{ for } i = 1,2,\ldots,k\}$.

Since $\mathbb{Q}$ is countable, the Cartesian product $A = \mathbb{Q}^k$ is countable.

To show that $A$ is dense in $\mathbb{R}^k$, let $x = (x_1, x_2, \ldots, x_k) \in \mathbb{R}^k$ and $\varepsilon > 0$. Since $\mathbb{Q}$ is dense in $\mathbb{R}$, for each $i$ there exists $q_i \in \mathbb{Q}$ such that $|x_i - q_i| < \varepsilon/\sqrt{k}$.

Then $q = (q_1, q_2, \ldots, q_k) \in A$ and $\|x - q\| = \sqrt{\sum_{i=1}^k (x_i - q_i)^2} < \sqrt{k(\varepsilon/\sqrt{k})^2} = \varepsilon$.

Therefore, $A$ is a countable dense subset of $\mathbb{R}^k$, so $\mathbb{R}^k$ is separable.

\begin{problembox}[3.34: Lindelöf Theorem in Separable Spaces]
Prove that the Lindelöf covering theorem (Theorem 3.28) is valid in any separable metric space.
\end{problembox}

\textbf{Solution:} Let $M$ be a separable metric space with countable dense subset $D = \{d_1, d_2, \ldots\}$. Let $\mathcal{F}$ be an open covering of $M$.

For each $d_i \in D$ and each positive rational $r$, if there exists a set $F \in \mathcal{F}$ such that $B(d_i;r) \subset F$, let $F_{i,r}$ be one such set.

The collection $\{F_{i,r} : i \in \mathbb{N}, r \in \mathbb{Q}^+, B(d_i;r) \subset F_{i,r} \text{ for some } F \in \mathcal{F}\}$ is countable.

We claim this collection covers $M$. Let $x \in M$. Since $\mathcal{F}$ covers $M$, there exists $F \in \mathcal{F}$ such that $x \in F$. Since $F$ is open, there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset F$.

Since $D$ is dense, there exists $d_i \in D$ such that $d_i \in B(x;\varepsilon/2)$. Let $r$ be a rational number such that $d(x,d_i) < r < \varepsilon/2$. Then $B(d_i;r) \subset B(x;\varepsilon) \subset F$.

Therefore, $F_{i,r}$ exists and contains $x$, showing that the countable subcollection covers $M$.

\begin{problembox}[3.35: Density and Open Sets]
If \( A \) is dense in \( S \) and if \( B \) is open in \( S \), prove that \( B \subseteq \overline{A \cap B} \).

\textit{Hint.} Exercise 3.13.
\end{problembox}

\textbf{Solution:}\\
The statement "$A$ is dense in $S$" means $S \subseteq \overline{A}$. We are given that $A \subseteq S$.
The statement "$B$ is open in $S$" means that $B = V \cap S$ for some set $V$ that is open in the larger metric space $M$.

Let $x \in B$. We want to show that $x \in \overline{A \cap B}$. This requires showing that any open neighborhood of $x$ in $M$ has a non-empty intersection with the set $A \cap B$.

Let $U$ be an arbitrary open neighborhood of $x$ in $M$.
Since $x \in B$ and $B = V \cap S$, we have $x \in V$.
The set $U \cap V$ is also an open neighborhood of $x$ because it is the intersection of two open sets.

Since $x \in B \subseteq S$ and $A$ is dense in $S$, $x$ is an adherent point of $A$. Therefore, the open neighborhood $U \cap V$ must contain a point from $A$. Let's call this point $y$.
So, $y \in (U \cap V) \cap A$.

Now we check if this point $y$ is in the required sets:
\begin{itemize}
    \item $y \in U$, so $y$ is in the arbitrary neighborhood of $x$.
    \item $y \in A$.
    \item We need to show $y \in B$. We know $y \in V$. Since we are given $A \subseteq S$, and $y \in A$, it follows that $y \in S$.
\end{itemize}
Since $y \in V$ and $y \in S$, we have $y \in V \cap S$, which means $y \in B$.

So we have found a point $y$ such that $y \in U$ and $y \in A \cap B$. This means $U \cap (A \cap B) \neq \emptyset$.
Since $U$ was an arbitrary open neighborhood of $x$, this proves that $x \in \overline{A \cap B}$.
As this holds for any $x \in B$, we conclude that $B \subseteq \overline{A \cap B}$.


\begin{problembox}[3.36: Intersection of Dense and Open Sets]
If each of \( A \) and \( B \) is dense in \( S \) and if \( B \) is open in \( S \), prove that \( A \cap B \) is dense in \( S \).
\end{problembox}

\textbf{Solution:} We need to show that $S \subseteq \overline{A \cap B}$.

Let $x \in S$. Since $B$ is open in $S$, there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \cap S \subset B$.

Since $A$ is dense in $S$, $B(x;\varepsilon) \cap A \neq \emptyset$. Let $y \in B(x;\varepsilon) \cap A$. Since $y \in S$ and $B(x;\varepsilon) \cap S \subset B$, we have $y \in B$.

Therefore, $y \in A \cap B$, so $B(x;\varepsilon) \cap (A \cap B) \neq \emptyset$.

This shows that every neighborhood of $x$ contains a point of $A \cap B$, so $x \in \overline{A \cap B}$.

Therefore, $S \subseteq \overline{A \cap B}$, showing that $A \cap B$ is dense in $S$.

\begin{problembox}[3.37: Product Metrics]
Given two metric spaces \((S_1, d_1)\) and \((S_2, d_2)\), a metric \( \rho \) for the Cartesian product \( S_1 \times S_2 \) can be constructed from \( d_1 \) and \( d_2 \) in many ways. For example, if \( x = (x_1, x_2) \) and \( y = (y_1, y_2) \) are in \( S_1 \times S_2 \), let \( \rho(x, y) = d_1(x_1, y_1) + d_2(x_2, y_2) \). Prove that \( \rho \) is a metric for \( S_1 \times S_2 \) and construct further examples.
\end{problembox}

\textbf{Solution:} We need to verify the three properties of a metric for $\rho(x,y) = d_1(x_1,y_1) + d_2(x_2,y_2)$:

(1) $\rho(x,y) \geq 0$ since $d_1(x_1,y_1) \geq 0$ and $d_2(x_2,y_2) \geq 0$.

(2) $\rho(x,y) = 0$ if and only if $d_1(x_1,y_1) = 0$ and $d_2(x_2,y_2) = 0$, which occurs if and only if $x_1 = y_1$ and $x_2 = y_2$, i.e., $x = y$.

(3) $\rho(x,y) = \rho(y,x)$ since $d_1(x_1,y_1) = d_1(y_1,x_1)$ and $d_2(x_2,y_2) = d_2(y_2,x_2)$.

(4) For the triangle inequality, let $z = (z_1,z_2)$. Then $\rho(x,z) = d_1(x_1,z_1) + d_2(x_2,z_2) \leq d_1(x_1,y_1) + d_1(y_1,z_1) + d_2(x_2,y_2) + d_2(y_2,z_2) = \rho(x,y) + \rho(y,z)$.

Other examples of product metrics include:
- $\rho(x,y) = \max\{d_1(x_1,y_1), d_2(x_2,y_2)\}$
- $\rho(x,y) = \sqrt{d_1(x_1,y_1)^2 + d_2(x_2,y_2)^2}$
- $\rho(x,y) = (d_1(x_1,y_1)^p + d_2(x_2,y_2)^p)^{1/p}$ for $p \geq 1$


\section{Compact subsets of a metric space}
Prove each of the following statements concerning an arbitrary metric space $(M,d)$ and subsets $S$, $T$ of $M$.

\begin{problembox}[3.38: Relative Compactness]
Assume \( S \subseteq T \subseteq M \). Then \( S \) is compact in \((M, d)\) if, and only if, \( S \) is compact in the metric subspace \((T, d)\).
\end{problembox}

\textbf{Solution:} Suppose $S$ is compact in $(M,d)$. Let $\mathcal{F}$ be an open covering of $S$ in the subspace $(T,d)$. Then each $F \in \mathcal{F}$ is of the form $F = U \cap T$ where $U$ is open in $(M,d)$.

The collection $\{U : U \text{ is open in } (M,d) \text{ and } U \cap T \in \mathcal{F}\}$ is an open covering of $S$ in $(M,d)$. Since $S$ is compact in $(M,d)$, there exists a finite subcollection $\{U_1, \ldots, U_n\}$ that covers $S$.

Then $\{U_1 \cap T, \ldots, U_n \cap T\}$ is a finite subcollection of $\mathcal{F}$ that covers $S$, showing that $S$ is compact in $(T,d)$.

Conversely, suppose $S$ is compact in $(T,d)$. Let $\mathcal{G}$ be an open covering of $S$ in $(M,d)$. Then $\{G \cap T : G \in \mathcal{G}\}$ is an open covering of $S$ in $(T,d)$. Since $S$ is compact in $(T,d)$, there exists a finite subcollection $\{G_1 \cap T, \ldots, G_n \cap T\}$ that covers $S$.

Then $\{G_1, \ldots, G_n\}$ is a finite subcollection of $\mathcal{G}$ that covers $S$, showing that $S$ is compact in $(M,d)$.

\begin{problembox}[3.39: Intersection with Compact Sets]
If \( S \) is closed and \( T \) is compact, then \( S \cap T \) is compact.
\end{problembox}

\textbf{Solution:} Since $T$ is compact, it is closed. Therefore, $S \cap T$ is the intersection of two closed sets, so it is closed.

Since $S \cap T \subseteq T$ and $T$ is compact, by Exercise 3.38, $S \cap T$ is compact in $(T,d)$. Since compactness is independent of the ambient space, $S \cap T$ is compact in $(M,d)$.

\begin{problembox}[3.40: Intersection of Compact Sets]
The intersection of an arbitrary collection of compact subsets of \( M \) is compact.
\end{problembox}

\textbf{Solution:} Let $\{K_\alpha\}$ be a collection of compact subsets of $M$. Since each $K_\alpha$ is closed, the intersection $\bigcap K_\alpha$ is closed.

Let $K_1$ be any member of the collection. Then $\bigcap K_\alpha \subseteq K_1$ and $K_1$ is compact. Since $\bigcap K_\alpha$ is closed and contained in a compact set, by Exercise 3.39, $\bigcap K_\alpha$ is compact.

\begin{problembox}[3.41: Finite Union of Compact Sets]
The union of a finite number of compact subsets of \( M \) is compact.
\end{problembox}

\textbf{Solution:} Let $K_1, K_2, \ldots, K_n$ be compact subsets of $M$. Since each $K_i$ is closed, their union $\bigcup_{i=1}^n K_i$ is closed.

Let $\mathcal{F}$ be an open covering of $\bigcup_{i=1}^n K_i$. Then $\mathcal{F}$ is also an open covering of each $K_i$. Since each $K_i$ is compact, there exists a finite subcollection $\mathcal{F}_i$ of $\mathcal{F}$ that covers $K_i$.

Then $\bigcup_{i=1}^n \mathcal{F}_i$ is a finite subcollection of $\mathcal{F}$ that covers $\bigcup_{i=1}^n K_i$.

Since $\bigcup_{i=1}^n K_i$ is closed and every open covering has a finite subcover, it is compact.

\begin{problembox}[3.42: Non-Compact Closed and Bounded Set]
Consider the metric space \( \mathbb{Q} \) of rational numbers with the Euclidean metric of \( \mathbb{R} \). Let \( S \) consist of all rational numbers in the open interval \((a, b)\), where \( a \) and \( b \) are irrational. Then \( S \) is a closed and bounded subset of \( \mathbb{Q} \) which is not compact.
\end{problembox}

\textbf{Solution:} Let $S = \mathbb{Q} \cap (a,b)$ where $a, b$ are irrational numbers.

$S$ is bounded since it is contained in the bounded interval $(a,b)$.

$S$ is closed in $\mathbb{Q}$ because its complement $\mathbb{Q} \setminus S = \mathbb{Q} \cap ((-\infty,a] \cup [b,\infty))$ is open in $\mathbb{Q}$.

However, $S$ is not compact. Let $\{q_n\}$ be a sequence of rational numbers in $(a,b)$ that converges to $a$ (which exists since $\mathbb{Q}$ is dense in $\mathbb{R}$). Then $\{q_n\}$ is a sequence in $S$ that has no convergent subsequence in $S$ (since $a \notin S$).

Therefore, $S$ is closed and bounded but not compact.

\begin{problembox}[Miscellaneous Properties of Interior and Boundary]
The following problems involve arbitrary subsets \( A \) and \( B \) of a metric space \( M \).
\end{problembox}

\section{Miscellaneous Properties of Interior and Boundary}
If $A$ and $B$ are subsets of a metric space $M$, prove that:

\begin{problembox}[3.43: Interior via Closure]
Prove that \(\text{int } A = M - \overline{M - A}\).
\end{problembox}

\textbf{Solution:} Let $x \in \text{int } A$. Then there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset A$. This means $B(x;\varepsilon) \cap (M - A) = \emptyset$, so $x \notin \overline{M - A}$. Therefore, $x \in M - \overline{M - A}$.

Conversely, let $x \in M - \overline{M - A}$. Then $x \notin \overline{M - A}$, so there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \cap (M - A) = \emptyset$. This means $B(x;\varepsilon) \subset A$, so $x \in \text{int } A$.

\begin{problembox}[3.44: Interior of Complement]
Prove that \(\text{int }(M - A) = M - \overline{A}\).
\end{problembox}

\textbf{Solution:} Let $x \in \text{int }(M - A)$. Then there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset M - A$. This means $B(x;\varepsilon) \cap A = \emptyset$, so $x \notin \overline{A}$. Therefore, $x \in M - \overline{A}$.

Conversely, let $x \in M - \overline{A}$. Then $x \notin \overline{A}$, so there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \cap A = \emptyset$. This means $B(x;\varepsilon) \subset M - A$, so $x \in \text{int }(M - A)$.

\begin{problembox}[3.45: Idempotence of Interior]
Prove that \(\text{int }(\text{int } A) = \text{int } A\).
\end{problembox}

\textbf{Solution:} Since $\text{int } A \subseteq A$, we have $\text{int }(\text{int } A) \subseteq \text{int } A$.

For the reverse inclusion, let $x \in \text{int } A$. Then there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset A$. Since $B(x;\varepsilon)$ is open and contained in $A$, we have $B(x;\varepsilon) \subset \text{int } A$. Therefore, $x \in \text{int }(\text{int } A)$.

\begin{problembox}[3.46: Interior of Intersections]
\begin{enumerate}[label=\alph*)]
\item Prove that \(\text{int } \left(\bigcap_{i=1}^n A_i\right) = \bigcap_{i=1}^n (\text{int } A_i)\), where each \( A_i \subseteq M \).
\item Show that \(\text{int } \left(\bigcap_{A \in F} A\right) \subseteq \bigcap_{A \in F} (\text{int } A)\) if \( F \) is an infinite collection of subsets of \( M \).
\item Give an example where equality does not hold in (b).
\end{enumerate}
\end{problembox}

\textbf{Solution:} 
(a) Let $x \in \text{int }(\bigcap_{i=1}^n A_i)$. Then there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset \bigcap_{i=1}^n A_i$. This means $B(x;\varepsilon) \subset A_i$ for each $i$, so $x \in \text{int } A_i$ for each $i$. Therefore, $x \in \bigcap_{i=1}^n (\text{int } A_i)$.

Conversely, let $x \in \bigcap_{i=1}^n (\text{int } A_i)$. Then for each $i$, there exists $\varepsilon_i > 0$ such that $B(x;\varepsilon_i) \subset A_i$. Let $\varepsilon = \min\{\varepsilon_1, \ldots, \varepsilon_n\}$. Then $B(x;\varepsilon) \subset \bigcap_{i=1}^n A_i$, so $x \in \text{int }(\bigcap_{i=1}^n A_i)$.

(b) Let $x \in \text{int }(\bigcap_{A \in F} A)$. Then there exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset \bigcap_{A \in F} A$. This means $B(x;\varepsilon) \subset A$ for each $A \in F$, so $x \in \text{int } A$ for each $A \in F$. Therefore, $x \in \bigcap_{A \in F} (\text{int } A)$.

(c) Let $F = \{A_n : n \in \mathbb{N}\}$ where $A_n = (-1/n, 1/n)$. Then $\bigcap_{A \in F} A = \{0\}$, so $\text{int }(\bigcap_{A \in F} A) = \emptyset$. However, $\text{int } A_n = A_n$ for each $n$, so $\bigcap_{A \in F} (\text{int } A) = \bigcap_{n=1}^{\infty} A_n = \{0\}$.

\begin{problembox}[3.47: Interior of Unions]
\begin{enumerate}[label=\alph*)]
\item Prove that \(\bigcup_{A \in F} (\text{int } A) \subseteq \text{int } \left(\bigcup_{A \in F} A\right)\).
\item Give an example of a finite collection \( F \) in which equality does not hold in (a).
\end{enumerate}
\end{problembox}

\textbf{Solution:} 
(a) Let $x \in \bigcup_{A \in F} (\text{int } A)$. Then $x \in \text{int } A$ for some $A \in F$. There exists $\varepsilon > 0$ such that $B(x;\varepsilon) \subset A \subset \bigcup_{A \in F} A$. Therefore, $x \in \text{int }(\bigcup_{A \in F} A)$.

(b) Let $F = \{A, B\}$ where $A = [0,1]$ and $B = [1,2]$. Then $\text{int } A = (0,1)$ and $\text{int } B = (1,2)$, so $\bigcup_{A \in F} (\text{int } A) = (0,1) \cup (1,2)$. However, $\bigcup_{A \in F} A = [0,2]$, so $\text{int }(\bigcup_{A \in F} A) = (0,2)$, which properly contains $(0,1) \cup (1,2)$.

\begin{problembox}[3.48: Interior of Boundary]
\begin{enumerate}[label=\alph*)]
\item Prove that \(\text{int } (\partial A) = \emptyset\) if \( A \) is open or if \( A \) is closed in \( M \).
\item Give an example in which \(\text{int } (\partial A) = M\).
\end{enumerate}
\end{problembox}

\textbf{Solution:} 
(a) If $A$ is open, then $\partial A = \overline{A} \setminus \text{int } A = \overline{A} \setminus A$. If $A$ is closed, then $\partial A = A \setminus \text{int } A$.

In both cases, $\partial A$ contains no open balls, so $\text{int } (\partial A) = \emptyset$.

(b) Let $A = \mathbb{Q}$ in the metric space $\mathbb{R}$. Then $\partial A = \mathbb{R}$, so $\text{int } (\partial A) = \mathbb{R} = M$.

\begin{problembox}[3.49: Interior of Union of Sets with Empty Interior]
If \(\text{int } A = \text{int } B = \emptyset\) and if \(A\) is closed in \(M\), then \(\text{int } (A \cup B) = \emptyset\).
\end{problembox}

\textbf{Solution:} Since $A$ is closed, $\text{int } A = \emptyset$ implies that $A$ has no isolated points. Therefore, every point in $A$ is a limit point of $A$.

Let $x \in A \cup B$. If $x \in A$, then every neighborhood of $x$ contains points of $A$ different from $x$. Since $A \subset A \cup B$, every neighborhood of $x$ contains points of $A \cup B$ different from $x$, so $x$ is not an interior point of $A \cup B$.

If $x \in B \setminus A$, then since $\text{int } B = \emptyset$, every neighborhood of $x$ contains points not in $B$. Since $A$ is closed and $x \notin A$, there exists a neighborhood of $x$ that doesn't intersect $A$. This neighborhood contains points not in $A \cup B$, so $x$ is not an interior point of $A \cup B$.

Therefore, $\text{int } (A \cup B) = \emptyset$.

\begin{problembox}[3.50: Counterexample for Union of Sets with Empty Interior]
Give an example in which \(\text{int } A = \text{int } B = \emptyset\) but \(\text{int } (A \cup B) = M\).
\end{problembox}

\textbf{Solution:} Let $A = \mathbb{Q}$ and $B = \mathbb{R} \setminus \mathbb{Q}$ in the metric space $\mathbb{R}$. Then $\text{int } A = \emptyset$ and $\text{int } B = \emptyset$, but $A \cup B = \mathbb{R}$, so $\text{int } (A \cup B) = \mathbb{R} = M$.

\begin{problembox}[3.51: Properties of Boundary]
Prove that:
\[
\partial A = \overline{A} \cap \overline{M - A} \quad \text{and} \quad \partial A = \partial(M - A).
\]
\end{problembox}

\textbf{Solution:} For the first equality, $x \in \partial A$ if and only if every neighborhood of $x$ contains both points of $A$ and points of $M - A$. This means $x \in \overline{A}$ and $x \in \overline{M - A}$, so $x \in \overline{A} \cap \overline{M - A}$.

For the second equality, $\partial A = \overline{A} \cap \overline{M - A} = \overline{M - A} \cap \overline{A} = \partial(M - A)$.

\begin{problembox}[3.52: Boundary of Union under Disjoint Closures]
If \(\overline{A} \cap \overline{B} = \emptyset\), then \(\partial(A \cup B) = \partial A \cup \partial B\).
\end{problembox}

\textbf{Solution:} Since $\overline{A} \cap \overline{B} = \emptyset$, we have $\overline{A \cup B} = \overline{A} \cup \overline{B}$.

Let $x \in \partial(A \cup B)$. Then $x \in \overline{A \cup B} = \overline{A} \cup \overline{B}$ and $x \in \overline{M - (A \cup B)} = \overline{(M - A) \cap (M - B)} \subseteq \overline{M - A} \cap \overline{M - B}$.

If $x \in \overline{A}$, then $x \in \overline{A} \cap \overline{M - A} = \partial A$. If $x \in \overline{B}$, then $x \in \overline{B} \cap \overline{M - B} = \partial B$. Therefore, $x \in \partial A \cup \partial B$.

Conversely, let $x \in \partial A \cup \partial B$. Without loss of generality, assume $x \in \partial A$. Then $x \in \overline{A} \subseteq \overline{A \cup B}$ and $x \in \overline{M - A} \subseteq \overline{M - (A \cup B)}$. Therefore, $x \in \partial(A \cup B)$.